---
title: "Final Prediction"
author: Ella Trembanis
date: '2024-11-03'
output:
  html_document:
    df_print: paged
categories: []
tags: []
slug: "final-prediction"
---

``` {r load libraries, echo=FALSE, include=FALSE, warning=FALSE}
library(car)
library(caret)
library(CVXR)
library(foreign)
library(glmnet)
library(haven)
library(janitor)
library(kableExtra)
library(maps)
library(mlr3)
library(randomForest)
library(ranger)
library(RColorBrewer)
library(sf)
library(tidyverse)
library(viridis)
library(cowplot)
library(curl)
library(geofacet)
library(randomForest)
library(rstan)
library(scales)
library(shinystan)
library(stargazer)
library(ggpubr)
library(ggthemes)
library(haven)
library(mgcv)
library(mgcViz)
library(spData)
library(tidygeocoder)
library(tigris)
library(tmap)
library(tmaptools)
library(knitr)
library(broom)
```

```{r custom theme, echo=FALSE, include=FALSE, warning=FALSE}
my_theme <- function() {
  theme(
  # set panel features
  panel.border = element_rect(color = "#000033", fill = NA, linetype = 1),
  panel.background = element_rect(fill = "#FFFFFF"),
  panel.grid.major.x = element_line(color = "#9999CC", linetype = 2, size = 0.2),
  panel.grid.minor.x = element_blank(),
  panel.grid.major.y = element_line(color = "#9999CC", linetype = 2, size = 0.2),
  panel.grid.minor.y = element_blank(),
  # set text and axis features
  text = element_text(color = "#000033", family = "Courier"),
  title = element_text(color = "#000033", face = "bold", family = "Courier"),
  axis.ticks = element_line(color = "#000033"),
  # set legend features
  legend.position = "bottom",
  legend.background = element_rect(color = "#000033", fill = NA, linetype = 1)
  )
}
```

```{r simplified custom theme, echo=FALSE, include=FALSE, warning=FALSE}
my_map_theme <- function() {
  theme(
    # set panel features
    panel.background = element_rect(fill = "#FFFFFF"),
    # set text features
    text = element_text(color = "#000033", family = "Courier"),
    title = element_text(color = "#000033", face = "bold", family = "Courier"),
    # set legend features
    legend.position = "bottom",
    legend.title.position = "top",
    legend.background = element_rect(color = "#000033", fill = NA, linetype = 1),
    # set axis features to not display
    axis.title = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank()
  )
}
```

```{r read data, echo=FALSE, include=FALSE, warning=FALSE}
# Read popular vote datasets. 
d_popvote <- read_csv("/Users/ellatrembanis/Desktop/Gov 1347/election-blog/election-blog/Final Prediction/data/popvote_1948_2020(2).csv")
d_state_popvote <- read_csv("/Users/ellatrembanis/Desktop/Gov 1347/election-blog/election-blog/Final Prediction/data/state_popvote_1948_2020(2).csv")

# Read elector distribution dataset. 
d_ec <- read_csv("/Users/ellatrembanis/Desktop/Gov 1347/election-blog/election-blog/Final Prediction/data/corrected_ec_1948_2024(2).csv")

# Read polling data. 
d_polls <- read_csv("/Users/ellatrembanis/Desktop/Gov 1347/election-blog/election-blog/Final Prediction/data/national_polls_1968-2024(7).csv")
d_state_polls <- read_csv("/Users/ellatrembanis/Desktop/Gov 1347/do not push/data/state_polls_1968-2024(7).csv")

# Process state-level polling data. 
d_pollav_state <- d_state_polls |> 
  group_by(year, state, party) |>
  mutate(mean_pollav = mean(poll_support, na.rm = TRUE)) |>
  top_n(1, poll_date) |> 
  rename(latest_pollav = poll_support) |>
  select(-c(weeks_left, days_left, poll_date, candidate, before_convention)) |>
  pivot_wider(names_from = party, values_from = c(latest_pollav, mean_pollav))

# Read economic data.
econ <- read_csv("/Users/ellatrembanis/Desktop/Gov 1347/election-blog/election-blog/Final Prediction/data/fred_econ.csv") |> 
# filter to most recent available quarter of 2024
  filter(quarter == 2)

# Read consumer sentiment index
ics <- read_csv("/Users/ellatrembanis/Desktop/Gov 1347/election-blog/election-blog/Final Prediction/data/tbqics.csv") |>
  mutate(
    quarter = case_when(QUARTER == "Jan.-Mar." ~ "Q1",
                        QUARTER == "Apr.-June" ~ "Q2",
                        QUARTER == "Jul.-Sep." ~ "Q3",
                        QUARTER == "Oct.-Dec." ~ "Q4")
  ) |>
  filter(quarter == "Q2") |>
  mutate(
    quarter = if_else(quarter == "Q2", 2, NA),
    year = YYYY,
    ics = ICS_ALL
    )

# Merge econ + ics
econ <- econ |>
  left_join(ics, by = join_by(year, quarter))

# Join data for time for change model.
tfc_train <- d_popvote |> 
  left_join(econ, by = "year") |> 
  filter(incumbent_party) |>
  mutate(incumbent = as.numeric(incumbent)) |>
  filter(year >= 1968)

# map of u.s. with state borders
states_map <- map_data("state")
```

```{r modelling, echo=FALSE, include=FALSE, warning=FALSE}
# use weeks left as columns
week_poll <- d_polls |> 
  group_by(year, party, weeks_left) |>
  summarize(mean_poll_week = mean(poll_support)) |> 
  filter(weeks_left <= 30) |> 
  pivot_wider(names_from = weeks_left, values_from = mean_poll_week) |>
# merge national vote
  left_join(d_popvote, by = c("year", "party"))

# split to train and test
week_poll_train <- week_poll |> 
  filter(year <= 2020)
week_poll_test <- week_poll |> 
  filter(year == 2024)

# rename
colnames(week_poll)[3:33] <- paste0("poll_weeks_left_", 0:30)
colnames(week_poll_train)[3:33] <- paste0("poll_weeks_left_", 0:30)
colnames(week_poll_test)[3:33] <- paste0("poll_weeks_left_", 0:30)

# merge
combined <- econ |> 
  left_join(week_poll, by = "year") |> 
  filter(year %in% c(unique(d_popvote$year), 2024)) |> 
  group_by(party) |> 
  mutate(pv2p_lag1 = lag(pv2p, 1), 
         pv2p_lag2 = lag(pv2p, 2)) |> 
  ungroup()

combined[29, "incumbent_party"] = TRUE
combined[30, "incumbent_party"] = FALSE
combined[29, "incumbent"] = FALSE

combined[1, "pv2p_lag1"] = 0.612033364797
combined[3, "pv2p_lag2"] = 0.612033364797
combined[2, "pv2p_lag1"] = 0.387966635203
combined[4, "pv2p_lag2"] = 0.387966635203
combined[1, "pv2p_lag2"] = 0.500874006373
combined[2, "pv2p_lag2"] = 0.499125993627

combined <- combined |> 
  mutate(gdp_growth_x_incumbent_party = GDP_growth_quarterly * incumbent_party,
         cpi_x_incumbent_party = CPI * incumbent_party)

combined_lags <- combined |>
  select(year, party, pv2p, pv2p_lag1, pv2p_lag2)

# merge datasets (polling + fundamentals + time for change)
combined_tfc <- tfc_train |>
  left_join(combined, by = join_by(year, candidate, GDP_growth_quarterly, party, pv, pv2p, incumbent, incumbent_party)) |>
  mutate(
    gdp_growth_x_prev_admin = GDP_growth_quarterly * prev_admin.x
  )

# add back in values lost in merge
combined_tfc[15, "incumbent"] = 0 # Harris incumbency status
combined_tfc[15, "poll_weeks_left_5"] = 48.52034 # Harris poll at t-minus 5 weeks
combined_tfc[15, "pv2p_lag1"] = 52.2698591 # Biden pv2p in 2020
combined_tfc[15, "gdp_growth_x_incumbent_party"] = 3.0 # gdp growth & party incumbency interaction term
combined_tfc[15, "ics"] = 71.1

# sequester 2024 data
combined_tfc_24 <- combined_tfc |>
  filter(year == 2024)
```

```{r states take 2, echo=FALSE, include=FALSE, warning=FALSE}
# popvote
d_popvote_simplified <- d_popvote |>
  select(year, party, candidate, incumbent, pv2p, prev_admin, incumbent_party) |>
  mutate(
    party = if_else(party == "republican", "REP", "DEM")
  )

# ics and gdp
d_econ_vars <- econ |>
  select(year, ics, GDP_growth_quarterly)

# nat poll
d_natpoll <- week_poll |>
  group_by(year, party) |>
  summarize(
    poll_week_1 = mean(poll_weeks_left_1)
  ) |>
  filter(year >= 1972)
 
# merging

d_popvote_simplified <- d_popvote_simplified |>
  left_join(d_econ_vars, by = "year") |>
  filter(year >= 1972) |>
  select(year, party, candidate, prev_admin, ics, pv2p, incumbent, incumbent_party, GDP_growth_quarterly)

d_popvote_simplified <- d_popvote_simplified |>
  left_join(d_natpoll) |>
  filter(year >= 1972) |>
  select(year, party, candidate, prev_admin, ics, pv2p, incumbent, incumbent_party, GDP_growth_quarterly, poll_week_1)
  
d_state_popvote_simplified <- d_state_popvote |>
  pivot_longer(
    cols = c(D_pv2p, R_pv2p),
    names_to = "party",
    values_to = "state_pv2p"
  ) |>
  mutate(
    party = if_else(party == "D_pv2p", "DEM", "REP")
  ) |>
  select(year, state, party, state_pv2p)

d_state_popvote_2024 <- data.frame(year = rep(2024, 102),
                                   state = rep(c(unique(d_state_popvote_simplified$state)), 2),
                                   party = rep(c(unique(d_state_popvote_simplified$party)), 51),
                                   state_pv2p = rep(NA, 102))

d_state_popvote_simplified <- rbind(d_state_popvote_simplified, d_state_popvote_2024)

# average state polls
d_state_polls_simplified <- d_state_polls |>
  group_by(state, year, party) |>
  summarize(
    mean_support = mean(poll_support, na.rm=T)
  )

#more  merging
d_state_tfmc <- d_state_popvote_simplified |>
  left_join(d_state_polls_simplified)

d_state_tfmc <- d_state_tfmc |>
  left_join(d_popvote_simplified, join_by(year, party)) |>
  filter(party == "DEM" & year >= "1972")

# create lag & poll error variables
d_state_tfmc <- d_state_tfmc |>
  mutate(
    state_poll_error = state_pv2p - poll_week_1
  )

# create tfmc dataframe for every state
for (variable in unique(d_state_tfmc$state)) {
  assign(variable, d_state_tfmc %>% filter (state == variable), envir = .GlobalEnv)
}

# separate 2024 and predict by state (where 2024 polls are available!)
Arizona <- Arizona |> mutate(state_pv2p_lag = lag(state_pv2p, n=1)) # populate lag variable
Arizona[1, "state_pv2p_lag"] = d_state_popvote$D_pv2p[d_state_popvote$year == 1968 & d_state_popvote$state == "Arizona"] # add back in 1972 lag
Arizona24 <- Arizona |> filter(year == 2024) # testing data
tfmc_az <- lm(state_pv2p ~ ics + mean_support + prev_admin + state_pv2p_lag + poll_week_1, data = Arizona) # regression
summary(tfmc_az) # produce summary
az_pred <- predict(tfmc_az, Arizona24, interval = "none") # point estimate of dem's popular vote share
Arizona20 <- Arizona |> filter(year == 2020) # 2020 testing data
Arizona16 <- Arizona |> filter(year == 2016) # 2016 testing data

California <- California |> mutate(state_pv2p_lag = lag(state_pv2p, n=1))
California[1, "state_pv2p_lag"] = d_state_popvote$D_pv2p[d_state_popvote$year == 1968 & d_state_popvote$state == "California"]
California$state_pv2p_lag[California$year == 2024] <- California$state_pv2p[California$year == 2020]
California24 <- California |> filter(year == 2024)
tfmc_ca <- lm(state_pv2p ~ ics + mean_support + prev_admin + state_pv2p_lag + poll_week_1, data = California)
summary(tfmc_ca)
ca_pred <- predict(tfmc_ca, California24, interval = "none")
California20 <- California |> filter(year == 2020)
California16 <- California |> filter(year == 2016)

Florida <- Florida |> mutate(state_pv2p_lag = lag(state_pv2p, n=1))
Florida[1, "state_pv2p_lag"] = d_state_popvote$D_pv2p[d_state_popvote$year == 1968 & d_state_popvote$state == "Florida"]
Florida$state_pv2p_lag[Florida$year == 2024] <- Florida$state_pv2p[Florida$year == 2020]
Florida24 <- Florida |> filter(year == 2024)
tfmc_fl <- lm(state_pv2p ~ ics + mean_support + prev_admin + state_pv2p_lag + poll_week_1, data = Florida)
summary(tfmc_fl)
fl_pred <- predict(tfmc_fl, Florida24, interval = "none")
Florida20 <- Florida |> filter(year == 2020)
Florida16 <- Florida |> filter(year == 2016)

Georgia <- Georgia |> mutate(state_pv2p_lag = lag(state_pv2p, n=1))
Georgia[1, "state_pv2p_lag"] = d_state_popvote$D_pv2p[d_state_popvote$year == 1968 & d_state_popvote$state == "Georgia"]
Georgia$state_pv2p_lag[Georgia$year == 2024] <- Georgia$state_pv2p[Georgia$year == 2020]
Georgia24 <- Georgia |> filter(year == 2024)
tfmc_ga <- lm(state_pv2p ~ ics + mean_support + prev_admin + state_pv2p_lag + poll_week_1, data = Georgia)
summary(tfmc_ga)
ga_pred <- predict(tfmc_ga, Georgia24, interval = "none")
Georgia20 <- Georgia |> filter(year == 2020)
Georgia16 <- Georgia |> filter(year == 2016)

Indiana <- Indiana |> mutate(state_pv2p_lag = lag(state_pv2p, n=1))
Indiana[1, "state_pv2p_lag"] = d_state_popvote$D_pv2p[d_state_popvote$year == 1968 & d_state_popvote$state == "Indiana"]
Indiana$state_pv2p_lag[Indiana$year == 2024] <- Indiana$state_pv2p[Indiana$year == 2020]
Indiana24 <- Indiana |> filter(year == 2024)
tfmc_in <- lm(state_pv2p ~ ics + mean_support + prev_admin + state_pv2p_lag + poll_week_1, data = Indiana)
summary(tfmc_in)
in_pred <- predict(tfmc_in, Indiana24, interval = "none")
Indiana20 <- Indiana |> filter(year == 2020)
Indiana16 <- Indiana |> filter(year == 2016)

Maryland <- Maryland |> mutate(state_pv2p_lag = lag(state_pv2p, n=1))
Maryland[1, "state_pv2p_lag"] = d_state_popvote$D_pv2p[d_state_popvote$year == 1968 & d_state_popvote$state == "Maryland"]
Maryland$state_pv2p_lag[Maryland$year == 2024] <- Maryland$state_pv2p[Maryland$year == 2020]
Maryland24 <- Maryland |> filter(year == 2024)
tfmc_md <- lm(state_pv2p ~ ics + mean_support + prev_admin + state_pv2p_lag + poll_week_1, data = Maryland)
summary(tfmc_md)
md_pred <- predict(tfmc_md, Maryland24, interval = "none")
Maryland20 <- Maryland |> filter(year == 2020)
Maryland16 <- Maryland |> filter(year == 2016)

Massachusetts <- Massachusetts |> mutate(state_pv2p_lag = lag(state_pv2p, n=1))
Massachusetts[1, "state_pv2p_lag"] = d_state_popvote$D_pv2p[d_state_popvote$year == 1968 & d_state_popvote$state == "Massachusetts"]
Massachusetts$state_pv2p_lag[Massachusetts$year == 2024] <- Massachusetts$state_pv2p[Massachusetts$year == 2020]
Massachusetts24 <- Massachusetts |> filter(year == 2024)
tfmc_ma <- lm(state_pv2p ~ ics + mean_support + prev_admin + state_pv2p_lag + poll_week_1, data = Massachusetts)
summary(tfmc_ma)
ma_pred <- predict(tfmc_ma, Massachusetts24, interval = "none")
Massachusetts20 <- Massachusetts |> filter(year == 2020)
Massachusetts16 <- Massachusetts |> filter(year == 2016)

Michigan <- Michigan |> mutate(state_pv2p_lag = lag(state_pv2p, n=1))
Michigan[1, "state_pv2p_lag"] = d_state_popvote$D_pv2p[d_state_popvote$year == 1968 & d_state_popvote$state == "Michigan"]
Michigan$state_pv2p_lag[Michigan$year == 2024] <- Michigan$state_pv2p[Michigan$year == 2020]
Michigan24 <- Michigan |> filter(year == 2024)
tfmc_mi <- lm(state_pv2p ~ ics + mean_support + prev_admin + state_pv2p_lag + poll_week_1, data = Michigan)
summary(tfmc_mi)
mi_pred <- predict(tfmc_mi, Michigan24, interval = "none")
Michigan20 <- Michigan |> filter(year == 2020)
Michigan16 <- Michigan |> filter(year == 2016)

Minnesota <- Minnesota |> mutate(state_pv2p_lag = lag(state_pv2p, n=1))
Minnesota[1, "state_pv2p_lag"] = d_state_popvote$D_pv2p[d_state_popvote$year == 1968 & d_state_popvote$state == "Minnesota"]
Minnesota$state_pv2p_lag[Minnesota$year == 2024] <- Minnesota$state_pv2p[Minnesota$year == 2020]
Minnesota24 <- Minnesota |> filter(year == 2024)
tfmc_mn <- lm(state_pv2p ~ ics + mean_support + prev_admin + state_pv2p_lag + poll_week_1, data = Minnesota)
summary(tfmc_mn)
mn_pred <- predict(tfmc_mn, Minnesota24, interval = "none")
Minnesota20 <- Minnesota |> filter(year == 2020)
Minnesota16 <- Minnesota |> filter(year == 2016)

Missouri <- Missouri |> mutate(state_pv2p_lag = lag(state_pv2p, n=1))
Missouri[1, "state_pv2p_lag"] = d_state_popvote$D_pv2p[d_state_popvote$year == 1968 & d_state_popvote$state == "Missouri"]
Missouri$state_pv2p_lag[Missouri$year == 2024] <- Missouri$state_pv2p[Missouri$year == 2020]
Missouri24 <- Missouri |> filter(year == 2024)
tfmc_mo <- lm(state_pv2p ~ ics + mean_support + prev_admin + state_pv2p_lag + poll_week_1, data = Missouri)
summary(tfmc_mo)
mo_pred <- predict(tfmc_mo, Missouri24, interval = "none")
Missouri20 <- Missouri |> filter(year == 2020)
Missouri16 <- Missouri |> filter(year == 2016)

Montana <- Montana |> mutate(state_pv2p_lag = lag(state_pv2p, n=1))
Montana[1, "state_pv2p_lag"] = d_state_popvote$D_pv2p[d_state_popvote$year == 1968 & d_state_popvote$state == "Montana"]
Montana$state_pv2p_lag[Montana$year == 2024] <- Montana$state_pv2p[Montana$year == 2020]
Montana24 <- Montana |> filter(year == 2024)
tfmc_mt <- lm(state_pv2p ~ ics + mean_support + prev_admin + state_pv2p_lag + poll_week_1, data = Montana)
summary(tfmc_mt)
mt_pred <- predict(tfmc_mt, Montana24, interval = "none")
Montana20 <- Montana |> filter(year == 2020)
Montana16 <- Montana |> filter(year == 2016)

Nebraska <- Nebraska |> mutate(state_pv2p_lag = lag(state_pv2p, n=1))
Nebraska[1, "state_pv2p_lag"] = d_state_popvote$D_pv2p[d_state_popvote$year == 1968 & d_state_popvote$state == "Nebraska"]
Nebraska$state_pv2p_lag[Nebraska$year == 2024] <- Nebraska$state_pv2p[Nebraska$year == 2020]
Nebraska24 <- Nebraska |> filter(year == 2024)
tfmc_ne <- lm(state_pv2p ~ ics + mean_support + prev_admin + state_pv2p_lag + poll_week_1, data = Nebraska)
summary(tfmc_ne)
ne_pred <- predict(tfmc_ne, Nebraska24, interval = "none")
Nebraska20 <- Nebraska |> filter(year == 2020)
Nebraska16 <- Nebraska |> filter(year == 2016)

Nevada <- Nevada |> mutate(state_pv2p_lag = lag(state_pv2p, n=1))
Nevada[1, "state_pv2p_lag"] = d_state_popvote$D_pv2p[d_state_popvote$year == 1968 & d_state_popvote$state == "Nevada"]
Nevada$state_pv2p_lag[Nevada$year == 2024] <- Nevada$state_pv2p[Nevada$year == 2020]
Nevada24 <- Nevada |> filter(year == 2024)
tfmc_nv <- lm(state_pv2p ~ ics + mean_support + prev_admin + state_pv2p_lag + poll_week_1, data = Nevada)
summary(tfmc_nv)
nv_pred <- predict(tfmc_nv, Nevada24, interval = "none")
Nevada20 <- Nevada |> filter(year == 2020)
Nevada16 <- Nevada |> filter(year == 2016)

`New Hampshire` <- `New Hampshire` |> mutate(state_pv2p_lag = lag(state_pv2p, n=1))
`New Hampshire`[1, "state_pv2p_lag"] = d_state_popvote$D_pv2p[d_state_popvote$year == 1968 & d_state_popvote$state == "New Hampshire"]
`New Hampshire`$state_pv2p_lag[`New Hampshire`$year == 2024] <- `New Hampshire`$state_pv2p[`New Hampshire`$year == 2020]
NH24 <- `New Hampshire` |> filter(year == 2024)
tfmc_nh <- lm(state_pv2p ~ ics + mean_support + prev_admin + state_pv2p_lag + poll_week_1, data = `New Hampshire`)
summary(tfmc_nh)
nh_pred <- predict(tfmc_nh, NH24, interval = "none")
NH20 <- `New Hampshire` |> filter(year == 2020)
NH16 <- `New Hampshire` |> filter(year == 2016)

`New Mexico` <- `New Mexico` |> mutate(state_pv2p_lag = lag(state_pv2p, n=1))
`New Mexico`[1, "state_pv2p_lag"] = d_state_popvote$D_pv2p[d_state_popvote$year == 1968 & d_state_popvote$state == "New Mexico"]
`New Mexico`$state_pv2p_lag[`New Mexico`$year == 2024] <- `New Mexico`$state_pv2p[`New Mexico`$year == 2020]
NM24 <- `New Mexico` |> filter(year == 2024)
tfmc_nm <- lm(state_pv2p ~ ics + mean_support + prev_admin + state_pv2p_lag + poll_week_1, data = `New Mexico`)
summary(tfmc_nm)
nm_pred <- predict(tfmc_nm, NM24, interval = "none")
NM20 <- `New Mexico` |> filter(year == 2020)
NM16 <- `New Mexico` |> filter(year == 2016)

`New York` <- `New York` |> mutate(state_pv2p_lag = lag(state_pv2p, n=1))
`New York`[1, "state_pv2p_lag"] = d_state_popvote$D_pv2p[d_state_popvote$year == 1968 & d_state_popvote$state == "New York"]
`New York`$state_pv2p_lag[`New York`$year == 2024] <- `New York`$state_pv2p[`New York`$year == 2020]
NY24 <- `New York` |> filter(year == 2024)
tfmc_ny <- lm(state_pv2p ~ ics + mean_support + prev_admin + state_pv2p_lag + poll_week_1, data = `New York`)
summary(tfmc_ny)
ny_pred <- predict(tfmc_ny, NY24, interval = "none")
NY20 <- `New York` |> filter(year == 2020)
NY16 <- `New York` |> filter(year == 2016)

`North Carolina` <- `North Carolina` |> mutate(state_pv2p_lag = lag(state_pv2p, n=1))
`North Carolina`[1, "state_pv2p_lag"] = d_state_popvote$D_pv2p[d_state_popvote$year == 1968 & d_state_popvote$state == "North Carolina"]
`North Carolina`$state_pv2p_lag[`North Carolina`$year == 2024] <- `North Carolina`$state_pv2p[`North Carolina`$year == 2020]
NC24 <- `North Carolina` |> filter(year == 2024)
tfmc_nc <- lm(state_pv2p ~ ics + mean_support + prev_admin + state_pv2p_lag + poll_week_1, data = `North Carolina`)
summary(tfmc_nc)
nc_pred <- predict(tfmc_nc, NC24, interval = "none")
NC20 <- `North Carolina` |> filter(year == 2020)
NC16 <- `North Carolina` |> filter(year == 2016)

Ohio <- Ohio |> mutate(state_pv2p_lag = lag(state_pv2p, n=1))
Ohio[1, "state_pv2p_lag"] = d_state_popvote$D_pv2p[d_state_popvote$year == 1968 & d_state_popvote$state == "Ohio"]
Ohio$state_pv2p_lag[Ohio$year == 2024] <- Ohio$state_pv2p[Ohio$year == 2020]
Ohio24 <- Ohio |> filter(year == 2024)
tfmc_oh <- lm(state_pv2p ~ ics + mean_support + prev_admin + state_pv2p_lag + poll_week_1, data = Ohio)
summary(tfmc_oh)
oh_pred <- predict(tfmc_oh, Ohio24, interval = "none")
Ohio20 <- Ohio |> filter(year == 2020)
Ohio16 <- Ohio |> filter(year == 2016)

Pennsylvania <- Pennsylvania |> mutate(state_pv2p_lag = lag(state_pv2p, n=1))
Pennsylvania[1, "state_pv2p_lag"] = d_state_popvote$D_pv2p[d_state_popvote$year == 1968 & d_state_popvote$state == "Pennsylvania"]
Pennsylvania$state_pv2p_lag[Pennsylvania$year == 2024] <- Pennsylvania$state_pv2p[Pennsylvania$year == 2020]
Pennsylvania24 <- Pennsylvania |> filter(year == 2024)
tfmc_pa <- lm(state_pv2p ~ ics + mean_support + prev_admin + state_pv2p_lag + poll_week_1, data = Pennsylvania)
summary(tfmc_pa)
pa_pred <- predict(tfmc_pa, Pennsylvania24, interval = "none")
Pennsylvania20 <- Pennsylvania |> filter(year == 2020)
Pennsylvania16 <- Pennsylvania |> filter(year == 2016)

Texas <- Texas |> mutate(state_pv2p_lag = lag(state_pv2p, n=1))
Texas[1, "state_pv2p_lag"] = d_state_popvote$D_pv2p[d_state_popvote$year == 1968 & d_state_popvote$state == "Texas"]
Texas$state_pv2p_lag[Texas$year == 2024] <- Texas$state_pv2p[Texas$year == 2020]
Texas24 <- Texas |> filter(year == 2024)
tfmc_tx <- lm(state_pv2p ~ ics + mean_support + prev_admin + state_pv2p_lag + poll_week_1, data = Texas)
summary(tfmc_tx)
tx_pred <- predict(tfmc_tx, Texas24, interval = "none")
Texas20 <- Texas |> filter(year == 2020)
Texas16 <- Texas |> filter(year == 2016)

Virginia <- Virginia |> mutate(state_pv2p_lag = lag(state_pv2p, n=1))
Virginia[1, "state_pv2p_lag"] = d_state_popvote$D_pv2p[d_state_popvote$year == 1968 & d_state_popvote$state == "Virginia"]
Virginia$state_pv2p_lag[Virginia$year == 2024] <- Virginia$state_pv2p[Virginia$year == 2020]
Virginia24 <- Virginia |> filter(year == 2024)
tfmc_va <- lm(state_pv2p ~ ics + mean_support + prev_admin + state_pv2p_lag + poll_week_1, data = Virginia)
summary(tfmc_va)
va_pred <- predict(tfmc_va, Virginia24, interval = "none")
Virginia20 <- Virginia |> filter(year == 2020)
Virginia16 <- Virginia |> filter(year == 2016)

Washington <- Washington |> mutate(state_pv2p_lag = lag(state_pv2p, n=1))
Washington[1, "state_pv2p_lag"] = d_state_popvote$D_pv2p[d_state_popvote$year == 1968 & d_state_popvote$state == "Washington"]
Washington$state_pv2p_lag[Washington$year == 2024] <- Washington$state_pv2p[Washington$year == 2020]
Washington24 <- Washington |> filter(year == 2024)
tfmc_wa <- lm(state_pv2p ~ ics + mean_support + prev_admin + state_pv2p_lag + poll_week_1, data = Washington)
summary(tfmc_wa)
wa_pred <- predict(tfmc_wa, Washington24, interval = "none")
Washington20 <- Washington |> filter(year == 2020)
Washington16 <- Washington |> filter(year == 2016)

Wisconsin <- Wisconsin |> mutate(state_pv2p_lag = lag(state_pv2p, n=1))
Wisconsin[1, "state_pv2p_lag"] = d_state_popvote$D_pv2p[d_state_popvote$year == 1968 & d_state_popvote$state == "Wisconsin"]
Wisconsin$state_pv2p_lag[Wisconsin$year == 2024] <- Wisconsin$state_pv2p[Wisconsin$year == 2020]
Wisconsin24 <- Wisconsin |> filter(year == 2024)
tfmc_wi <- lm(state_pv2p ~ ics + mean_support + prev_admin + state_pv2p_lag + poll_week_1, data = Wisconsin)
summary(tfmc_wi)
wi_pred <- predict(tfmc_wi, Wisconsin24, interval = "none")
Wisconsin20 <- Wisconsin |> filter(year == 2020)
Wisconsin16 <- Wisconsin |> filter(year == 2016)
```

```{r states no polls, echo=FALSE, include=FALSE, warning = FALSE}
# separate 2024 and predict by state (where 2024 polls are not available!)
Alabama <- Alabama |> mutate(state_pv2p_lag = lag(state_pv2p, n=1))
Alabama[1, "state_pv2p_lag"] = d_state_popvote$D_pv2p[d_state_popvote$year == 1968 & d_state_popvote$state == "Alabama"]
Alabama24 <- data.frame(year = 2024,
                        state = "Alabama",
                        party = "DEM",
                        state_pv2p = NA,
                        ics = unique(d_state_tfmc$ics[d_state_tfmc$year == 2024]),
                        prev_admin = TRUE,
                        poll_week_1 = unique(d_state_tfmc$poll_week_1[d_state_tfmc$year == 2024]),
                        state_pv2p_lag = Alabama$state_pv2p[Alabama$year == 2020],
                        poll_error = Alabama$state_poll_error[Alabama$year == 2020],
                        mean_support = Alabama$poll_week_1[Alabama$year == 2024] + Alabama$state_poll_error[Alabama$year == 2020])
tfmc_al <- lm(state_pv2p ~ ics + mean_support + prev_admin + state_pv2p_lag + poll_week_1, data = Alabama)
summary(tfmc_al)
al_pred <- predict(tfmc_al, Alabama24, interval = "none")
Alabama20 <- Alabama |> filter(year == 2020)
Alabama16 <- Alabama |> filter(year == 2016)

Alaska <- Alaska |> mutate(state_pv2p_lag = lag(state_pv2p, n=1))
Alaska[1, "state_pv2p_lag"] = d_state_popvote$D_pv2p[d_state_popvote$year == 1968 & d_state_popvote$state == "Alaska"]
Alaska24 <- data.frame(year = 2024,
                        state = "Alaska",
                        party = "DEM",
                        state_pv2p = NA,
                        ics = unique(d_state_tfmc$ics[d_state_tfmc$year == 2024]),
                        prev_admin = TRUE,
                        poll_week_1 = unique(d_state_tfmc$poll_week_1[d_state_tfmc$year == 2024]),
                        state_pv2p_lag = Alaska$state_pv2p[Alaska$year == 2020],
                        poll_error = Alaska$state_poll_error[Alaska$year == 2020],
                        mean_support = Alaska$poll_week_1[Alaska$year == 2024] + Alaska$state_poll_error[Alaska$year == 2020])
tfmc_ak <- lm(state_pv2p ~ ics + mean_support + prev_admin + state_pv2p_lag + poll_week_1, data = Alaska)
summary(tfmc_ak)
ak_pred <- predict(tfmc_ak, Alaska24, interval = "none")
Alaska20 <- Alaska |> filter(year == 2020)
Alaska16 <- Alaska |> filter(year == 2016)

Arkansas <- Arkansas |> mutate(state_pv2p_lag = lag(state_pv2p, n=1))
Arkansas[1, "state_pv2p_lag"] = d_state_popvote$D_pv2p[d_state_popvote$year == 1968 & d_state_popvote$state == "Arkansas"]
Arkansas24 <- data.frame(year = 2024,
                        state = "Arkansas",
                        party = "DEM",
                        state_pv2p = NA,
                        ics = unique(d_state_tfmc$ics[d_state_tfmc$year == 2024]),
                        prev_admin = TRUE,
                        poll_week_1 = unique(d_state_tfmc$poll_week_1[d_state_tfmc$year == 2024]),
                        state_pv2p_lag = Arkansas$state_pv2p[Arkansas$year == 2020],
                        poll_error = Arkansas$state_poll_error[Arkansas$year == 2020],
                        mean_support = Arkansas$poll_week_1[Arkansas$year == 2024] + Arkansas$state_poll_error[Arkansas$year == 2020])
tfmc_ar <- lm(state_pv2p ~ ics + mean_support + prev_admin + state_pv2p_lag + poll_week_1, data = Arkansas)
summary(tfmc_ar)
ar_pred <- predict(tfmc_ar, Arkansas24, interval = "none")
Arkansas20 <- Arkansas |> filter(year == 2020)
Arkansas16 <- Arkansas |> filter(year == 2016)

Colorado <- Colorado |> mutate(state_pv2p_lag = lag(state_pv2p, n=1))
Colorado[1, "state_pv2p_lag"] = d_state_popvote$D_pv2p[d_state_popvote$year == 1968 & d_state_popvote$state == "Colorado"]
Colorado24 <- data.frame(year = 2024,
                        state = "Colorado",
                        party = "DEM",
                        state_pv2p = NA,
                        ics = unique(d_state_tfmc$ics[d_state_tfmc$year == 2024]),
                        prev_admin = TRUE,
                        poll_week_1 = unique(d_state_tfmc$poll_week_1[d_state_tfmc$year == 2024]),
                        state_pv2p_lag = Colorado$state_pv2p[Colorado$year == 2020],
                        poll_error = Colorado$state_poll_error[Colorado$year == 2020],
                        mean_support = Colorado$poll_week_1[Colorado$year == 2024] + Colorado$state_poll_error[Colorado$year == 2020])
tfmc_co <- lm(state_pv2p ~ ics + mean_support + prev_admin + state_pv2p_lag + poll_week_1, data = Colorado)
summary(tfmc_co)
co_pred <- predict(tfmc_co, Colorado24, interval = "none")
Colorado20 <- Colorado |> filter(year == 2020)
Colorado16 <- Colorado |> filter(year == 2016)

Connecticut <- Connecticut |> mutate(state_pv2p_lag = lag(state_pv2p, n=1))
Connecticut[1, "state_pv2p_lag"] = d_state_popvote$D_pv2p[d_state_popvote$year == 1968 & d_state_popvote$state == "Connecticut"]
Connecticut24 <- data.frame(year = 2024,
                        state = "Connecticut",
                        party = "DEM",
                        state_pv2p = NA,
                        ics = unique(d_state_tfmc$ics[d_state_tfmc$year == 2024]),
                        prev_admin = TRUE,
                        poll_week_1 = unique(d_state_tfmc$poll_week_1[d_state_tfmc$year == 2024]),
                        state_pv2p_lag = Connecticut$state_pv2p[Connecticut$year == 2020],
                        poll_error = Connecticut$state_poll_error[Connecticut$year == 2020],
                        mean_support = Connecticut$poll_week_1[Connecticut$year == 2024] + Connecticut$state_poll_error[Connecticut$year == 2020])
tfmc_ct <- lm(state_pv2p ~ ics + mean_support + prev_admin + state_pv2p_lag + poll_week_1, data = Connecticut)
summary(tfmc_ct)
ct_pred <- predict(tfmc_ct, Connecticut24, interval = "none")
Connecticut20 <- Connecticut |> filter(year == 2020)
Connecticut16 <- Connecticut |> filter(year == 2016)

Delaware <- Delaware |> mutate(state_pv2p_lag = lag(state_pv2p, n=1))
Delaware[1, "state_pv2p_lag"] = d_state_popvote$D_pv2p[d_state_popvote$year == 1968 & d_state_popvote$state == "Delaware"]
Delaware24 <- data.frame(year = 2024,
                        state = "Delaware",
                        party = "DEM",
                        state_pv2p = NA,
                        ics = unique(d_state_tfmc$ics[d_state_tfmc$year == 2024]),
                        prev_admin = TRUE,
                        poll_week_1 = unique(d_state_tfmc$poll_week_1[d_state_tfmc$year == 2024]),
                        state_pv2p_lag = Delaware$state_pv2p[Delaware$year == 2020],
                        poll_error = Delaware$state_poll_error[Delaware$year == 2020],
                        mean_support = Delaware$poll_week_1[Delaware$year == 2024] + Delaware$state_poll_error[Delaware$year == 2020])
tfmc_de <- lm(state_pv2p ~ ics + mean_support + prev_admin + state_pv2p_lag + poll_week_1, data = Delaware)
summary(tfmc_de)
de_pred <- predict(tfmc_de, Delaware24, interval = "none")
Delaware20 <- Delaware |> filter(year == 2020)
Delaware16 <- Delaware |> filter(year == 2016)

Hawaii <- Hawaii |> mutate(state_pv2p_lag = lag(state_pv2p, n=1))
Hawaii[1, "state_pv2p_lag"] = d_state_popvote$D_pv2p[d_state_popvote$year == 1968 & d_state_popvote$state == "Hawaii"]
Hawaii24 <- data.frame(year = 2024,
                        state = "Hawaii",
                        party = "DEM",
                        state_pv2p = NA,
                        ics = unique(d_state_tfmc$ics[d_state_tfmc$year == 2024]),
                        prev_admin = TRUE,
                        poll_week_1 = unique(d_state_tfmc$poll_week_1[d_state_tfmc$year == 2024]),
                        state_pv2p_lag = Hawaii$state_pv2p[Hawaii$year == 2020],
                        poll_error = Hawaii$state_poll_error[Hawaii$year == 2020],
                        mean_support = Hawaii$poll_week_1[Hawaii$year == 2024] + Hawaii$state_poll_error[Hawaii$year == 2020])
tfmc_hi <- lm(state_pv2p ~ ics + mean_support + prev_admin + state_pv2p_lag + poll_week_1, data = Hawaii)
summary(tfmc_hi)
hi_pred <- predict(tfmc_hi, Hawaii24, interval = "none")
Hawaii20 <- Hawaii |> filter(year == 2020)
Hawaii16 <- Hawaii |> filter(year == 2016)

Idaho <- Idaho |> mutate(state_pv2p_lag = lag(state_pv2p, n=1))
Idaho[1, "state_pv2p_lag"] = d_state_popvote$D_pv2p[d_state_popvote$year == 1968 & d_state_popvote$state == "Idaho"]
Idaho24 <- data.frame(year = 2024,
                        state = "Idaho",
                        party = "DEM",
                        state_pv2p = NA,
                        ics = unique(d_state_tfmc$ics[d_state_tfmc$year == 2024]),
                        prev_admin = TRUE,
                        poll_week_1 = unique(d_state_tfmc$poll_week_1[d_state_tfmc$year == 2024]),
                        state_pv2p_lag = Idaho$state_pv2p[Idaho$year == 2020],
                        poll_error = Idaho$state_poll_error[Idaho$year == 2020],
                        mean_support = Idaho$poll_week_1[Idaho$year == 2024] + Idaho$state_poll_error[Idaho$year == 2020])
tfmc_id <- lm(state_pv2p ~ ics + mean_support + prev_admin + state_pv2p_lag + poll_week_1, data = Idaho)
summary(tfmc_id)
id_pred <- predict(tfmc_id, Idaho24, interval = "none")
Idaho20 <- Idaho |> filter(year == 2020)
Idaho16 <- Idaho |> filter(year == 2016)

Illinois <- Illinois |> mutate(state_pv2p_lag = lag(state_pv2p, n=1))
Illinois[1, "state_pv2p_lag"] = d_state_popvote$D_pv2p[d_state_popvote$year == 1968 & d_state_popvote$state == "Illinois"]
Illinois24 <- data.frame(year = 2024,
                        state = "Illinois",
                        party = "DEM",
                        state_pv2p = NA,
                        ics = unique(d_state_tfmc$ics[d_state_tfmc$year == 2024]),
                        prev_admin = TRUE,
                        poll_week_1 = unique(d_state_tfmc$poll_week_1[d_state_tfmc$year == 2024]),
                        state_pv2p_lag = Illinois$state_pv2p[Illinois$year == 2020],
                        poll_error = Illinois$state_poll_error[Illinois$year == 2020],
                        mean_support = Illinois$poll_week_1[Illinois$year == 2024] + Illinois$state_poll_error[Illinois$year == 2020])
tfmc_il <- lm(state_pv2p ~ ics + mean_support + prev_admin + state_pv2p_lag + poll_week_1, data = Illinois)
summary(tfmc_il)
il_pred <- predict(tfmc_il, Illinois24, interval = "none")
Illinois20 <- Illinois |> filter(year == 2020)
Illinois16 <- Illinois |> filter(year == 2016)

Iowa <- Iowa |> mutate(state_pv2p_lag = lag(state_pv2p, n=1))
Iowa[1, "state_pv2p_lag"] = d_state_popvote$D_pv2p[d_state_popvote$year == 1968 & d_state_popvote$state == "Iowa"]
Iowa24 <- data.frame(year = 2024,
                     state = "Iowa",
                     party = "DEM",
                     state_pv2p = NA,
                     ics = unique(d_state_tfmc$ics[d_state_tfmc$year == 2024]),
                     prev_admin = TRUE,
                     poll_week_1 = unique(d_state_tfmc$poll_week_1[d_state_tfmc$year == 2024]),
                     state_pv2p_lag = Iowa$state_pv2p[Iowa$year == 2020],
                     poll_error = Iowa$state_poll_error[Iowa$year == 2020],
                     mean_support = Iowa$poll_week_1[Iowa$year == 2024] + Iowa$state_poll_error[Iowa$year == 2020])
tfmc_ia <- lm(state_pv2p ~ ics + mean_support + prev_admin + state_pv2p_lag + poll_week_1, data = Iowa)
summary(tfmc_ia)
ia_pred <- predict(tfmc_ia, Iowa24, interval = "none")
Iowa20 <- Iowa |> filter(year == 2020)
Iowa16 <- Iowa |> filter(year == 2016)

Kansas <- Kansas |> mutate(state_pv2p_lag = lag(state_pv2p, n=1))
Kansas[1, "state_pv2p_lag"] = d_state_popvote$D_pv2p[d_state_popvote$year == 1968 & d_state_popvote$state == "Kansas"]
Kansas24 <- data.frame(year = 2024,
                       state = "Kansas",
                       party = "DEM",
                       state_pv2p = NA,
                       ics = unique(d_state_tfmc$ics[d_state_tfmc$year == 2024]),
                       prev_admin = TRUE,
                       poll_week_1 = unique(d_state_tfmc$poll_week_1[d_state_tfmc$year == 2024]),
                       state_pv2p_lag = Kansas$state_pv2p[Kansas$year == 2020],
                       poll_error = Kansas$state_poll_error[Kansas$year == 2020],
                       mean_support = Kansas$poll_week_1[Kansas$year == 2024] + Kansas$state_poll_error[Kansas$year == 2020])
tfmc_ks <- lm(state_pv2p ~ ics + mean_support + prev_admin + state_pv2p_lag + poll_week_1, data = Kansas)
summary(tfmc_ks)
ks_pred <- predict(tfmc_ks, Kansas24, interval = "none")
Kansas20 <- Kansas |> filter(year == 2020)
Kansas16 <- Kansas |> filter(year == 2016)

Kentucky <- Kentucky |> mutate(state_pv2p_lag = lag(state_pv2p, n=1))
Kentucky[1, "state_pv2p_lag"] = d_state_popvote$D_pv2p[d_state_popvote$year == 1968 & d_state_popvote$state == "Kentucky"]
Kentucky24 <- data.frame(year = 2024,
                         state = "Kentucky",
                         party = "DEM",
                         state_pv2p = NA,
                         ics = unique(d_state_tfmc$ics[d_state_tfmc$year == 2024]),
                         prev_admin = TRUE,
                         poll_week_1 = unique(d_state_tfmc$poll_week_1[d_state_tfmc$year == 2024]),
                         state_pv2p_lag = Kentucky$state_pv2p[Kentucky$year == 2020],
                         poll_error = Kentucky$state_poll_error[Kentucky$year == 2020],
                         mean_support = Kentucky$poll_week_1[Kentucky$year == 2024] + Kentucky$state_poll_error[Kentucky$year == 2020])
tfmc_ky <- lm(state_pv2p ~ ics + mean_support + prev_admin + state_pv2p_lag + poll_week_1, data = Kentucky)
summary(tfmc_ky)
ky_pred <- predict(tfmc_ky, Kentucky24, interval = "none")
Kentucky20 <- Kentucky |> filter(year == 2020)
Kentucky16 <- Kentucky |> filter(year == 2016)

Louisiana <- Louisiana |> mutate(state_pv2p_lag = lag(state_pv2p, n=1))
Louisiana[1, "state_pv2p_lag"] = d_state_popvote$D_pv2p[d_state_popvote$year == 1968 & d_state_popvote$state == "Louisiana"]
Louisiana24 <- data.frame(year = 2024,
                          state = "Louisiana",
                          party = "DEM",
                          state_pv2p = NA,
                          ics = unique(d_state_tfmc$ics[d_state_tfmc$year == 2024]),
                          prev_admin = TRUE,
                          poll_week_1 = unique(d_state_tfmc$poll_week_1[d_state_tfmc$year == 2024]),
                          state_pv2p_lag = Louisiana$state_pv2p[Louisiana$year == 2020],
                          poll_error = Louisiana$state_poll_error[Louisiana$year == 2020],
                          mean_support = Louisiana$poll_week_1[Louisiana$year == 2024] + Louisiana$state_poll_error[Louisiana$year == 2020])
tfmc_la <- lm(state_pv2p ~ ics + mean_support + prev_admin + state_pv2p_lag + poll_week_1, data = Louisiana)
summary(tfmc_la)
la_pred <- predict(tfmc_la, Louisiana24, interval = "none")
Louisiana20 <- Louisiana |> filter(year == 2020)
Louisiana16 <- Louisiana |> filter(year == 2016)

Maine <- Maine |> mutate(state_pv2p_lag = lag(state_pv2p, n=1))
Maine[1, "state_pv2p_lag"] = d_state_popvote$D_pv2p[d_state_popvote$year == 1968 & d_state_popvote$state == "Maine"]
Maine24 <- data.frame(year = 2024,
                      state = "Maine",
                      party = "DEM",
                      state_pv2p = NA,
                      ics = unique(d_state_tfmc$ics[d_state_tfmc$year == 2024]),
                      prev_admin = TRUE,
                      poll_week_1 = unique(d_state_tfmc$poll_week_1[d_state_tfmc$year == 2024]),
                      state_pv2p_lag = Maine$state_pv2p[Maine$year == 2020],
                      poll_error = Maine$state_poll_error[Maine$year == 2020],
                      mean_support = Maine$poll_week_1[Maine$year == 2024] + Maine$state_poll_error[Maine$year == 2020])
tfmc_me <- lm(state_pv2p ~ ics + mean_support + prev_admin + state_pv2p_lag + poll_week_1, data = Maine)
summary(tfmc_me)
me_pred <- predict(tfmc_me, Maine24, interval = "none")
Maine20 <- Maine |> filter(year == 2020)
Maine16 <- Maine |> filter(year == 2016)

Mississippi <- Mississippi |> mutate(state_pv2p_lag = lag(state_pv2p, n=1))
Mississippi[1, "state_pv2p_lag"] = d_state_popvote$D_pv2p[d_state_popvote$year == 1968 & d_state_popvote$state == "Mississippi"]
Mississippi24 <- data.frame(year = 2024,
                            state = "Mississippi",
                            party = "DEM",
                            state_pv2p = NA,
                            ics = unique(d_state_tfmc$ics[d_state_tfmc$year == 2024]),
                            prev_admin = TRUE,
                            poll_week_1 = unique(d_state_tfmc$poll_week_1[d_state_tfmc$year == 2024]),
                            state_pv2p_lag = Mississippi$state_pv2p[Mississippi$year == 2020],
                            poll_error = Mississippi$state_poll_error[Mississippi$year == 2020],
                            mean_support = Mississippi$poll_week_1[Mississippi$year == 2024] + Mississippi$state_poll_error[Mississippi$year == 2020])
tfmc_ms <- lm(state_pv2p ~ ics + mean_support + prev_admin + state_pv2p_lag + poll_week_1, data = Mississippi)
summary(tfmc_ms)
ms_pred <- predict(tfmc_ms, Mississippi24, interval = "none")
Mississippi20 <- Mississippi |> filter(year == 2020)
Mississippi16 <- Mississippi |> filter(year == 2016)

`New Jersey` <- `New Jersey` |> mutate(state_pv2p_lag = lag(state_pv2p, n=1))
`New Jersey`[1, "state_pv2p_lag"] = d_state_popvote$D_pv2p[d_state_popvote$year == 1968 & d_state_popvote$state == "New Jersey"]
NJ24 <- data.frame(year = 2024,
                           state = "New Jersey",
                           party = "DEM",
                           state_pv2p = NA,
                           ics = unique(d_state_tfmc$ics[d_state_tfmc$year == 2024]),
                           prev_admin = TRUE,
                           poll_week_1 = unique(d_state_tfmc$poll_week_1[d_state_tfmc$year == 2024]),
                           state_pv2p_lag = `New Jersey`$state_pv2p[`New Jersey`$year == 2020],
                           poll_error = `New Jersey`$state_poll_error[`New Jersey`$year == 2020],
                           mean_support = `New Jersey`$poll_week_1[`New Jersey`$year == 2024] + `New Jersey`$state_poll_error[`New Jersey`$year == 2020])
tfmc_nj <- lm(state_pv2p ~ ics + mean_support + prev_admin + state_pv2p_lag + poll_week_1, data = `New Jersey`)
summary(tfmc_nj)
nj_pred <- predict(tfmc_nj, NJ24, interval = "none")
NJ20 <- `New Jersey` |> filter(year == 2020)
NJ16 <- `New Jersey` |> filter(year == 2016)

`North Dakota` <- `North Dakota` |> mutate(state_pv2p_lag = lag(state_pv2p, n=1))
`North Dakota`[1, "state_pv2p_lag"] = d_state_popvote$D_pv2p[d_state_popvote$year == 1968 & d_state_popvote$state == "North Dakota"]
ND24 <- data.frame(year = 2024,
                             state = "North Dakota",
                             party = "DEM",
                             state_pv2p = NA,
                             ics = unique(d_state_tfmc$ics[d_state_tfmc$year == 2024]),
                             prev_admin = TRUE,
                             poll_week_1 = unique(d_state_tfmc$poll_week_1[d_state_tfmc$year == 2024]),
                             state_pv2p_lag = `North Dakota`$state_pv2p[`North Dakota`$year == 2020],
                             poll_error = `North Dakota`$state_poll_error[`North Dakota`$year == 2020],
                             mean_support = `North Dakota`$poll_week_1[`North Dakota`$year == 2024] + `North Dakota`$state_poll_error[`North Dakota`$year == 2020])
tfmc_nd <- lm(state_pv2p ~ ics + mean_support + prev_admin + state_pv2p_lag + poll_week_1, data = `North Dakota`)
summary(tfmc_nd)
nd_pred <- predict(tfmc_nd, ND24, interval = "none")
ND20 <- `North Dakota` |> filter(year == 2020)
ND16 <- `North Dakota` |> filter(year == 2016)

Oklahoma <- Oklahoma |> mutate(state_pv2p_lag = lag(state_pv2p, n=1))
Oklahoma[1, "state_pv2p_lag"] = d_state_popvote$D_pv2p[d_state_popvote$year == 1968 & d_state_popvote$state == "Oklahoma"]
Oklahoma24 <- data.frame(year = 2024,
                         state = "Oklahoma",
                         party = "DEM",
                         state_pv2p = NA,
                         ics = unique(d_state_tfmc$ics[d_state_tfmc$year == 2024]),
                         prev_admin = TRUE,
                        poll_week_1 = unique(d_state_tfmc$poll_week_1[d_state_tfmc$year == 2024]),
                         state_pv2p_lag = Oklahoma$state_pv2p[Oklahoma$year == 2020],
                         poll_error = Oklahoma$state_poll_error[Oklahoma$year == 2020],
                         mean_support = Oklahoma$poll_week_1[Oklahoma$year == 2024] + Oklahoma$state_poll_error[Oklahoma$year == 2020])
tfmc_ok <- lm(state_pv2p ~ ics + mean_support + prev_admin + state_pv2p_lag + poll_week_1, data = Oklahoma)
summary(tfmc_ok)
ok_pred <- predict(tfmc_ok, Oklahoma24, interval = "none")
Oklahoma20 <- Oklahoma |> filter(year == 2020)
Oklahoma16 <- Oklahoma |> filter(year == 2016)

Oregon <- Oregon |> mutate(state_pv2p_lag = lag(state_pv2p, n=1))
Oregon[1, "state_pv2p_lag"] = d_state_popvote$D_pv2p[d_state_popvote$year == 1968 & d_state_popvote$state == "Oregon"]
Oregon24 <- data.frame(year = 2024,
                       state = "Oregon",
                       party = "DEM",
                       state_pv2p = NA,
                       ics = unique(d_state_tfmc$ics[d_state_tfmc$year == 2024]),
                       prev_admin = TRUE,
                       poll_week_1 = unique(d_state_tfmc$poll_week_1[d_state_tfmc$year == 2024]),
                       state_pv2p_lag = Oregon$state_pv2p[Oregon$year == 2020],
                       poll_error = Oregon$state_poll_error[Oregon$year == 2020],
                       mean_support = Oregon$poll_week_1[Oregon$year == 2024] + Oregon$state_poll_error[Oregon$year == 2020])
tfmc_or <- lm(state_pv2p ~ ics + mean_support + prev_admin + state_pv2p_lag + poll_week_1, data = Oregon)
summary(tfmc_or)
or_pred <- predict(tfmc_or, Oregon24, interval = "none")
Oregon20 <- Oregon |> filter(year == 2020)
Oregon16 <- Oregon |> filter(year == 2016)

`Rhode Island` <- `Rhode Island` |> mutate(state_pv2p_lag = lag(state_pv2p, n=1))
`Rhode Island`[1, "state_pv2p_lag"] = d_state_popvote$D_pv2p[d_state_popvote$year == 1968 & d_state_popvote$state == "Rhode Island"]
RI24 <- data.frame(year = 2024,
                             state = "Rhode Island",
                             party = "DEM",
                             state_pv2p = NA,
                             ics = unique(d_state_tfmc$ics[d_state_tfmc$year == 2024]),
                             prev_admin = TRUE,
                             poll_week_1 = unique(d_state_tfmc$poll_week_1[d_state_tfmc$year == 2024]),
                             state_pv2p_lag = `Rhode Island`$state_pv2p[`Rhode Island`$year == 2020],
                             poll_error = `Rhode Island`$state_poll_error[`Rhode Island`$year == 2020],
                             mean_support = `Rhode Island`$poll_week_1[`Rhode Island`$year == 2024] + `Rhode Island`$state_poll_error[`Rhode Island`$year == 2020])
tfmc_ri <- lm(state_pv2p ~ ics + mean_support + prev_admin + state_pv2p_lag + poll_week_1, data = `Rhode Island`)
summary(tfmc_ri)
ri_pred <- predict(tfmc_ri, RI24, interval = "none")
RI20 <- `Rhode Island` |> filter(year == 2020)
RI16 <- `Rhode Island` |> filter(year == 2016)

`South Carolina` <- `South Carolina` |> mutate(state_pv2p_lag = lag(state_pv2p, n=1))
`South Carolina`[1, "state_pv2p_lag"] = d_state_popvote$D_pv2p[d_state_popvote$year == 1968 & d_state_popvote$state == "South Carolina"]
SC24 <- data.frame(year = 2024,
                               state = "South Carolina",
                               party = "DEM",
                               state_pv2p = NA,
                               ics = unique(d_state_tfmc$ics[d_state_tfmc$year == 2024]),
                               prev_admin = TRUE,
                               poll_week_1 = unique(d_state_tfmc$poll_week_1[d_state_tfmc$year == 2024]),
                               state_pv2p_lag = `South Carolina`$state_pv2p[`South Carolina`$year == 2020],
                               poll_error = `South Carolina`$state_poll_error[`South Carolina`$year == 2020],
                               mean_support = `South Carolina`$poll_week_1[`South Carolina`$year == 2024] + `South Carolina`$state_poll_error[`South Carolina`$year == 2020])
tfmc_sc <- lm(state_pv2p ~ ics + mean_support + prev_admin + state_pv2p_lag + poll_week_1, data = `South Carolina`)
summary(tfmc_sc)
sc_pred <- predict(tfmc_sc, SC24, interval = "none")
SC20 <- `South Carolina` |> filter(year == 2020)
SC16 <- `South Carolina` |> filter(year == 2016)

`South Dakota` <- `South Dakota` |> mutate(state_pv2p_lag = lag(state_pv2p, n=1))
`South Dakota`[1, "state_pv2p_lag"] = d_state_popvote$D_pv2p[d_state_popvote$year == 1968 & d_state_popvote$state == "South Dakota"]
SD24 <- data.frame(year = 2024,
                             state = "South Dakota",
                             party = "DEM",
                             state_pv2p = NA,
                             ics = unique(d_state_tfmc$ics[d_state_tfmc$year == 2024]),
                             prev_admin = TRUE,
                             poll_week_1 = unique(d_state_tfmc$poll_week_1[d_state_tfmc$year == 2024]),
                             state_pv2p_lag = `South Dakota`$state_pv2p[`South Dakota`$year == 2020],
                             poll_error = `South Dakota`$state_poll_error[`South Dakota`$year == 2020],
                             mean_support = `South Dakota`$poll_week_1[`South Dakota`$year == 2024] + `South Dakota`$state_poll_error[`South Dakota`$year == 2020])
tfmc_sd <- lm(state_pv2p ~ ics + mean_support + prev_admin + state_pv2p_lag + poll_week_1, data = `South Dakota`)
summary(tfmc_sd)
sd_pred <- predict(tfmc_sd, SD24, interval = "none")
SD20 <- `South Dakota` |> filter(year == 2020)
SD16 <- `South Dakota` |> filter(year == 2016)

Tennessee <- Tennessee |> mutate(state_pv2p_lag = lag(state_pv2p, n=1))
Tennessee[1, "state_pv2p_lag"] = d_state_popvote$D_pv2p[d_state_popvote$year == 1968 & d_state_popvote$state == "Tennessee"]
Tennessee24 <- data.frame(year = 2024,
                          state = "Tennessee",
                          party = "DEM",
                          state_pv2p = NA,
                          ics = unique(d_state_tfmc$ics[d_state_tfmc$year == 2024]),
                          prev_admin = TRUE,
                          poll_week_1 = unique(d_state_tfmc$poll_week_1[d_state_tfmc$year == 2024]),
                          state_pv2p_lag = Tennessee$state_pv2p[Tennessee$year == 2020],
                          poll_error = Tennessee$state_poll_error[Tennessee$year == 2020],
                          mean_support = Tennessee$poll_week_1[Tennessee$year == 2024] + Tennessee$state_poll_error[Tennessee$year == 2020])
tfmc_tn <- lm(state_pv2p ~ ics + mean_support + prev_admin + state_pv2p_lag + poll_week_1, data = Tennessee)
summary(tfmc_tn)
tn_pred <- predict(tfmc_tn, Tennessee24, interval = "none")
Tennessee20 <- Tennessee |> filter(year == 2020)
Tennessee16 <- Tennessee |> filter(year == 2016)

Utah <- Utah |> mutate(state_pv2p_lag = lag(state_pv2p, n=1))
Utah[1, "state_pv2p_lag"] = d_state_popvote$D_pv2p[d_state_popvote$year == 1968 & d_state_popvote$state == "Utah"]
Utah24 <- data.frame(year = 2024,
                     state = "Utah",
                     party = "DEM",
                     state_pv2p = NA,
                     ics = unique(d_state_tfmc$ics[d_state_tfmc$year == 2024]),
                     prev_admin = TRUE,
                     poll_week_1 = unique(d_state_tfmc$poll_week_1[d_state_tfmc$year == 2024]),
                     state_pv2p_lag = Utah$state_pv2p[Utah$year == 2020],
                     poll_error = Utah$state_poll_error[Utah$year == 2020],
                     mean_support = Utah$poll_week_1[Utah$year == 2024] + Utah$state_poll_error[Utah$year == 2020])
tfmc_ut <- lm(state_pv2p ~ ics + mean_support + prev_admin + state_pv2p_lag + poll_week_1, data = Utah)
summary(tfmc_ut)
ut_pred <- predict(tfmc_ut, Utah24, interval = "none")
Utah20 <- Utah |> filter(year == 2020)
Utah16 <- Utah |> filter(year == 2016)

Vermont <- Vermont |> mutate(state_pv2p_lag = lag(state_pv2p, n=1))
Vermont[1, "state_pv2p_lag"] = d_state_popvote$D_pv2p[d_state_popvote$year == 1968 & d_state_popvote$state == "Vermont"]
Vermont24 <- data.frame(year = 2024,
                        state = "Vermont",
                        party = "DEM",
                        state_pv2p = NA,
                        ics = unique(d_state_tfmc$ics[d_state_tfmc$year == 2024]),
                        prev_admin = TRUE,
                        poll_week_1 = unique(d_state_tfmc$poll_week_1[d_state_tfmc$year == 2024]),
                        state_pv2p_lag = Vermont$state_pv2p[Vermont$year == 2020],
                        poll_error = Vermont$state_poll_error[Vermont$year == 2020],
                        mean_support = Vermont$poll_week_1[Vermont$year == 2024] + Vermont$state_poll_error[Vermont$year == 2020])
tfmc_vt <- lm(state_pv2p ~ ics + mean_support + prev_admin + state_pv2p_lag + poll_week_1, data = Vermont)
summary(tfmc_vt)
vt_pred <- predict(tfmc_vt, Vermont24, interval = "none")
Vermont20 <- Vermont |> filter(year == 2020)
Vermont16 <- Vermont |> filter(year == 2016)

`West Virginia` <- `West Virginia` |> mutate(state_pv2p_lag = lag(state_pv2p, n=1))
`West Virginia`[1, "state_pv2p_lag"] = d_state_popvote$D_pv2p[d_state_popvote$year == 1968 & d_state_popvote$state == "West Virginia"]
WV24 <- data.frame(year = 2024,
                              state = "West Virginia",
                              party = "DEM",
                              state_pv2p = NA,
                              ics = unique(d_state_tfmc$ics[d_state_tfmc$year == 2024]),
                              prev_admin = TRUE,
                              poll_week_1 = unique(d_state_tfmc$poll_week_1[d_state_tfmc$year == 2024]),
                              state_pv2p_lag = `West Virginia`$state_pv2p[`West Virginia`$year == 2020],
                              poll_error = `West Virginia`$state_poll_error[`West Virginia`$year == 2020],
                              mean_support = `West Virginia`$poll_week_1[`West Virginia`$year == 2024] + `West Virginia`$state_poll_error[`West Virginia`$year == 2020])
tfmc_wv <- lm(state_pv2p ~ ics + mean_support + prev_admin + state_pv2p_lag + poll_week_1, data = `West Virginia`)
summary(tfmc_wv)
wv_pred <- predict(tfmc_wv, WV24, interval = "none")
WV20 <- `West Virginia` |> filter(year == 2020)
WV16 <- `West Virginia` |> filter(year == 2016)

Wyoming <- Wyoming |> mutate(state_pv2p_lag = lag(state_pv2p, n=1))
Wyoming[1, "state_pv2p_lag"] = d_state_popvote$D_pv2p[d_state_popvote$year == 1968 & d_state_popvote$state == "Wyoming"]
Wyoming24 <- data.frame(year = 2024,
                        state = "Wyoming",
                        party = "DEM",
                        state_pv2p = NA,
                        ics = unique(d_state_tfmc$ics[d_state_tfmc$year == 2024]),
                        prev_admin = TRUE,
                        poll_week_1 = unique(d_state_tfmc$poll_week_1[d_state_tfmc$year == 2024]),
                        state_pv2p_lag = Wyoming$state_pv2p[Wyoming$year == 2020],
                        poll_error = Wyoming$state_poll_error[Wyoming$year == 2020],
                        mean_support = Wyoming$poll_week_1[Wyoming$year == 2024] + Wyoming$state_poll_error[Wyoming$year == 2020])
tfmc_wy <- lm(state_pv2p ~ ics + mean_support + prev_admin + state_pv2p_lag + poll_week_1, data = Wyoming)
summary(tfmc_wy)
wy_pred <- predict(tfmc_wy, Wyoming24, interval = "none")
Wyoming20 <- Wyoming |> filter(year == 2020)
Wyoming16 <- Wyoming |> filter(year == 2016)

# Districts
DC_polls <- d_state_polls_simplified |> filter(state == "District of Columbia" & party == "DEM") |> mutate(state = if_else(state == "District of Columbia", "District Of Columbia", NA))
DC <- d_state_tfmc |> filter(state == "District Of Columbia") |>
  select(-mean_support, -state_poll_error)
DC <- DC |> left_join(DC_polls, by = join_by(year, state, party))
DC <- DC |> mutate(state_poll_error = state_pv2p - poll_week_1)
DC <- DC |> mutate(state_pv2p_lag = lag(state_pv2p, n=1))
DC[1, "state_pv2p_lag"] = d_state_popvote$D_pv2p[d_state_popvote$year == 1968 & d_state_popvote$state == "District Of Columbia"]
DC24 <- data.frame(year = 2024,
                        state = "District Of Columbia",
                        party = "DEM",
                        state_pv2p = NA,
                        ics = unique(d_state_tfmc$ics[d_state_tfmc$year == 2024]),
                        prev_admin = TRUE,
                        poll_week_1 = unique(d_state_tfmc$poll_week_1[d_state_tfmc$year == 2024]),
                        state_pv2p_lag = DC$state_pv2p[DC$year == 2020],
                        poll_error = DC$state_poll_error[DC$year == 2020],
                        mean_support = DC$poll_week_1[DC$year == 2024] + DC$state_poll_error[DC$year == 2020])
tfmc_dc <- lm(state_pv2p ~ ics + mean_support + prev_admin + state_pv2p_lag + poll_week_1, data = DC)
summary(tfmc_dc)
dc_pred <- predict(tfmc_dc, DC24, interval = "prediction", level = 0.8)
DC20 <- DC |> filter(year == 2020)
DC16 <- DC |> filter(year == 2016)

unpic_table <- tibble(Region = c("Alaska", "District Of Columbia", "Hawaii"),
                `Dem Two-Party Vote Share` = c(42.53449, 94.12381, 71.76807),
                Winner = c("Republican", "Democrat", "Democrat"))
table3 <- knitr::kable(unpic_table, digits = 2)
```

```{r pop vote, echo=FALSE, include=FALSE, warning=FALSE}
# Popular Vote Prediction
d_pop_tfmc <- d_state_tfmc |>
  filter(state == "Delaware") |> # filter to arbitrary state
  select(year, party, candidate, pv2p, ics, prev_admin, prev_admin, incumbent_party, GDP_growth_quarterly, poll_week_1) |> # choose only national vars
  mutate(pv2p_lag = lag(pv2p, n = 1))
d_pop_tfmc[1, "pv2p_lag"] = 0.493511945857
pop24 <- d_pop_tfmc |> filter(year == 2024) # testing data
tfmc_pop <- lm(pv2p ~ ics + poll_week_1 + prev_admin + pv2p_lag, data = d_pop_tfmc)
summary(tfmc_pop) # adjusted r-squared 0.7618
pop_pred <- predict(tfmc_pop, pop24, interval = "prediction", level = 0.8) # 51.26% Harris 2-party popular vote

table <- tibble(Estimate = c("Dem Two-Party Vote Share", "Lower Bound", "Upper Bound"),
                pv2p = c(51.25541, 46.51746, 55.99336))
table <- table |> pivot_wider(names_from = Estimate, values_from = pv2p)
table1 <- knitr::kable(table, digits = 2)

ics_outliers <- ggplot(d_pop_tfmc, aes(x=ics)) +
  geom_boxplot(fill = "#9999CC") +
  xlab("Second Quarter Consumer Sentiment") +
  ggtitle("Fig. I - Q2 ICS Boxplot") +
  my_theme() +
  theme(plot.title = element_text(size = 18),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())
```

```{r simulation, echo=FALSE, include=FALSE, warning=FALSE}
# Standardize number of columns - training data.
Alabama <- Alabama |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Alaska <- Alaska |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Arkansas <- Arkansas |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Arizona <- Arizona |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
California <- California |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Colorado <- Colorado |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Connecticut <- Connecticut |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Delaware <- Delaware |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Florida <- Florida |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Georgia <- Georgia |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Hawaii <- Hawaii |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Idaho <- Idaho |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Illinois <- Illinois |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Indiana <- Indiana |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Iowa <- Iowa |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Kansas <- Kansas |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Kentucky <- Kentucky |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Louisiana <- Louisiana |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Maine <- Maine |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Maryland <- Maryland |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Massachusetts <- Massachusetts |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Michigan <- Michigan |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Minnesota <- Minnesota |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Mississippi <- Mississippi |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Missouri <- Missouri |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Montana <- Montana |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Nebraska <- Nebraska |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Nevada <- Nevada |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
`New Hampshire` <- `New Hampshire` |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
`New Jersey` <- `New Jersey` |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
`New Mexico` <- `New Mexico` |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
`New York` <- `New York` |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
`North Carolina` <- `North Carolina` |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
`North Dakota` <- `North Dakota` |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Ohio <- Ohio |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Oklahoma <- Oklahoma |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Oregon <- Oregon |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Pennsylvania <- Pennsylvania |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
`Rhode Island` <- `Rhode Island` |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
`South Carolina` <- `South Carolina` |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
`South Dakota` <- `South Dakota` |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Tennessee <- Tennessee |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Texas <- Texas |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Utah <- Utah |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Vermont <- Vermont |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Virginia <- Virginia |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Washington <- Washington |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
`West Virginia` <- `West Virginia` |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Wisconsin <- Wisconsin |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Wyoming <- Wyoming |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)

# Subset to training data (pre-2024).
d_train <- rbind(Alabama, Alaska, Arizona, Arkansas, California, Colorado, Connecticut, Delaware, Florida, Georgia, Hawaii, Idaho, Illinois, Indiana, Iowa, Kansas, Kentucky, Louisiana, Maine, Maryland, Massachusetts, Michigan, Minnesota, Mississippi, Missouri, Montana, Nebraska, Nevada, `New Hampshire`, `New Jersey`, `New Mexico`, `New York`, `North Carolina`, `North Dakota`, Ohio, Oklahoma, Oregon, Pennsylvania, `Rhode Island`, `South Carolina`, `South Dakota`, Tennessee, Texas, Utah, Vermont, Virginia, Washington, `West Virginia`, Wisconsin, Wyoming) |>
  filter(year <= 2020)

# Subset to mock training data (pre-2020)
d_train20 <- rbind(Alabama, Alaska, Arizona, Arkansas, California, Colorado, Connecticut, Delaware, Florida, Georgia, Hawaii, Idaho, Illinois, Indiana, Iowa, Kansas, Kentucky, Louisiana, Maine, Maryland, Massachusetts, Michigan, Minnesota, Mississippi, Missouri, Montana, Nebraska, Nevada, `New Hampshire`, `New Jersey`, `New Mexico`, `New York`, `North Carolina`, `North Dakota`, Ohio, Oklahoma, Oregon, Pennsylvania, `Rhode Island`, `South Carolina`, `South Dakota`, Tennessee, Texas, Utah, Vermont, Virginia, Washington, `West Virginia`, Wisconsin, Wyoming) |>
  filter(year <= 2016)

# Subset to mock training data (pre-2016)
d_train16 <- rbind(Alabama, Alaska, Arizona, Arkansas, California, Colorado, Connecticut, Delaware, Florida, Georgia, Hawaii, Idaho, Illinois, Indiana, Iowa, Kansas, Kentucky, Louisiana, Maine, Maryland, Massachusetts, Michigan, Minnesota, Mississippi, Missouri, Montana, Nebraska, Nevada, `New Hampshire`, `New Jersey`, `New Mexico`, `New York`, `North Carolina`, `North Dakota`, Ohio, Oklahoma, Oregon, Pennsylvania, `Rhode Island`, `South Carolina`, `South Dakota`, Tennessee, Texas, Utah, Vermont, Virginia, Washington, `West Virginia`, Wisconsin, Wyoming) |>
  filter(year <= 2016)

# Standardize number of columns - testing data.
Alabama24 <- Alabama24 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Alaska24 <- Alaska24 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Arkansas24 <- Arkansas24 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Arizona24 <- Arizona24 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
California24 <- California24 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Colorado24 <- Colorado24 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Connecticut24 <- Connecticut24 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Delaware24 <- Delaware24 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Florida24 <- Florida24 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Georgia24 <- Georgia24 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Hawaii24 <- Hawaii24 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Idaho24 <- Idaho24 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Illinois24 <- Illinois24 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Indiana24 <- Indiana24 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Iowa24 <- Iowa24 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Kansas24 <- Kansas24 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Kentucky24 <- Kentucky24 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Louisiana24 <- Louisiana24 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Maine24 <- Maine24 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Maryland24 <- Maryland24 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Massachusetts24 <- Massachusetts24 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Michigan24 <- Michigan24 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Minnesota24 <- Minnesota24 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Mississippi24 <- Mississippi24 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Missouri24 <- Missouri24 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Montana24 <- Montana24 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Nebraska24 <- Nebraska24 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Nevada24 <- Nevada24 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
NH24 <- NH24 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
NJ24 <- NJ24 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
NM24 <- NM24 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
NY24 <- NY24 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
NC24 <- NC24 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
ND24 <- ND24 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Ohio24 <- Ohio24 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Oklahoma24 <- Oklahoma24 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Oregon24 <- Oregon24 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Pennsylvania24 <- Pennsylvania24 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
RI24 <- RI24 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
SC24 <- SC24 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
SD24 <- SD24 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Tennessee24 <- Tennessee24 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Texas24 <- Texas24 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Utah24 <- Utah24 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Vermont24 <- Vermont24 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Virginia24 <- Virginia24 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Washington24 <- Washington24 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
WV24 <- WV24 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Wisconsin24 <- Wisconsin24 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Wyoming24 <- Wyoming24 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)

Alabama20 <- Alabama20 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Alaska20 <- Alaska20 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Arkansas20 <- Arkansas20 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Arizona20 <- Arizona20 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
California20 <- California20 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Colorado20 <- Colorado20 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Connecticut20 <- Connecticut20 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Delaware20 <- Delaware20 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Florida20 <- Florida20 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Georgia20 <- Georgia20 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Hawaii20 <- Hawaii20 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Idaho20 <- Idaho20 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Illinois20 <- Illinois20 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Indiana20 <- Indiana20 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Iowa20 <- Iowa20 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Kansas20 <- Kansas20 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Kentucky20 <- Kentucky20 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Louisiana20 <- Louisiana20 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Maine20 <- Maine20 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Maryland20 <- Maryland20 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Massachusetts20 <- Massachusetts20 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Michigan20 <- Michigan20 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Minnesota20 <- Minnesota20 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Mississippi20 <- Mississippi20 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Missouri20 <- Missouri20 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Montana20 <- Montana20 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Nebraska20 <- Nebraska20 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Nevada20 <- Nevada20 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
NH20 <- NH20 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
NJ20 <- NJ20 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
NM20 <- NM20 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
NY20 <- NY20 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
NC20 <- NC20 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
ND20 <- ND20 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Ohio20 <- Ohio20 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Oklahoma20 <- Oklahoma20 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Oregon20 <- Oregon20 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Pennsylvania20 <- Pennsylvania20 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
RI20 <- RI20 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
SC20 <- SC20 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
SD20 <- SD20 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Tennessee20 <- Tennessee20 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Texas20 <- Texas20 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Utah20 <- Utah20 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Vermont20 <- Vermont20 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Virginia20 <- Virginia20 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Washington20 <- Washington20 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
WV20 <- WV20 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Wisconsin20 <- Wisconsin20 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Wyoming20 <- Wyoming20 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)

Alabama16 <- Alabama16 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Alaska16 <- Alaska16 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Arkansas16 <- Arkansas16 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Arizona16 <- Arizona16 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
California16 <- California16 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Colorado16 <- Colorado16 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Connecticut16 <- Connecticut16 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Delaware16 <- Delaware16 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Florida16 <- Florida16 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Georgia16 <- Georgia16 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Hawaii16 <- Hawaii16 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Idaho16 <- Idaho16 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Illinois16 <- Illinois16 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Indiana16 <- Indiana16 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Iowa16 <- Iowa16 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Kansas16 <- Kansas16 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Kentucky16 <- Kentucky16 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Louisiana16 <- Louisiana16 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Maine16 <- Maine16 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Maryland16 <- Maryland16 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Massachusetts16 <- Massachusetts16 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Michigan16 <- Michigan16 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Minnesota16 <- Minnesota16 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Mississippi16 <- Mississippi16 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Missouri16 <- Missouri16 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Montana16 <- Montana16 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Nebraska16 <- Nebraska16 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Nevada16 <- Nevada16 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
NH16 <- NH16 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
NJ16 <- NJ16 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
NM16 <- NM16 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
NY16 <- NY16 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
NC16 <- NC16 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
ND16 <- ND16 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Ohio16 <- Ohio16 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Oklahoma16 <- Oklahoma16 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Oregon16 <- Oregon16 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Pennsylvania16 <- Pennsylvania16 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
RI16 <- RI16 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
SC16 <- SC16 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
SD16 <- SD16 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Tennessee16 <- Tennessee16 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Texas16 <- Texas16 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Utah16 <- Utah16 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Vermont16 <- Vermont16 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Virginia16 <- Virginia16 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Washington16 <- Washington16 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
WV16 <- WV16 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Wisconsin16 <- Wisconsin16 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)
Wyoming16 <- Wyoming16 |> select(year, state, party, state_pv2p, ics, prev_admin, poll_week_1, state_pv2p_lag, mean_support)

# Subset to testing data (2024).
d_test <- rbind(Alabama24, Alaska24, Arizona24, Arkansas24, California24, Colorado24, Connecticut24, Delaware24, Florida24, Georgia24, Hawaii24, Idaho24, Illinois24, Indiana24, Iowa24, Kansas24, Kentucky24, Louisiana24, Maine24, Maryland24, Massachusetts24, Michigan24, Minnesota24, Mississippi24, Missouri24, Montana24, Nebraska24, Nevada24, NH24, NJ24, NM24, NY24, NC24, ND24, Ohio24, Oklahoma24, Oregon24, Pennsylvania24, RI24, SC24, SD24, Tennessee24, Texas24, Utah24, Vermont24, Virginia24, Washington24, WV24, Wisconsin24, Wyoming24)

# Subset to testing data (2020).
d_test20 <- rbind(Alabama20, Alaska20, Arizona20, Arkansas20, California20, Colorado20, Connecticut20, Delaware20, Florida20, Georgia20, Hawaii20, Idaho20, Illinois20, Indiana20, Iowa20, Kansas20, Kentucky20, Louisiana20, Maine20, Maryland20, Massachusetts20, Michigan20, Minnesota20, Mississippi20, Missouri20, Montana20, Nebraska20, Nevada20, NH20, NJ20, NM20, NY20, NC20, ND20, Ohio20, Oklahoma20, Oregon20, Pennsylvania20, RI20, SC20, SD20, Tennessee20, Texas20, Utah20, Vermont20, Virginia20, Washington20, WV20, Wisconsin20, Wyoming20)

# Subset to testing data (2016).
d_test16 <- rbind(Alabama16, Alaska16, Arizona16, Arkansas16, California16, Colorado16, Connecticut16, Delaware16, Florida16, Georgia16, Hawaii16, Idaho16, Illinois16, Indiana16, Iowa16, Kansas16, Kentucky16, Louisiana16, Maine16, Maryland16, Massachusetts16, Michigan16, Minnesota16, Mississippi16, Missouri16, Montana16, Nebraska16, Nevada16, NH16, NJ16, NM16, NY16, NC16, ND16, Ohio16, Oklahoma16, Oregon16, Pennsylvania16, RI16, SC16, SD16, Tennessee16, Texas16, Utah16, Vermont16, Virginia16, Washington16, WV16, Wisconsin16, Wyoming16)

# Run model. 
mod_lm_dem <- lm(state_pv2p ~ ics + prev_admin + poll_week_1 + state_pv2p_lag + mean_support, 
                 data = d_train)
summary(mod_lm_dem)

# Simple simulation. 
simp.vars <- c("ics", "prev_admin", "poll_week_1", "state_pv2p_lag", "mean_support")
mod_lm_dem_simp <- lm(state_pv2p ~ ics + prev_admin + poll_week_1 + state_pv2p_lag + mean_support,
                      data = d_train)
summary(mod_lm_dem_simp) # adjusted r-squared = 0.7796, all vars significant

# Subset testing data to only relevant variables for our simple model. 
d_test_simp <- d_test |>
  select(state, year, all_of(simp.vars))

# Predict.
simp_pred_dem <- predict(mod_lm_dem_simp, d_test_simp)
simp_pred_dem_interval <- predict(mod_lm_dem_simp, d_test_simp, interval = "prediction", level = 0.8)
simp_pred_dem_interval <- as.data.frame(simp_pred_dem_interval)
s

# Create dataset to summarize winners and EC vote distributions. 
win_pred <- data.frame(state = d_test_simp$state,
                       year = rep(2024, length(d_test_simp$state)),
                       simp_pred_dem = simp_pred_dem,
                       winner = ifelse(simp_pred_dem > 50, "Democrat", "Republican")) |>
  left_join(d_ec, by = c("state", "year")) |>
  mutate(
    region = tolower(state)
  )

win_pred <- win_pred |> 
  mutate(
    harris_win = ifelse(winner == "Democrat", 1, 0),
    harris_ec = (harris_win*electors)
  )
sum(win_pred$harris_ec) #223 Harris EC (not counting DC)

win_pred <- win_pred |> cbind(lower = simp_pred_dem_interval$lwr, upper = simp_pred_dem_interval$upr)

ec24_pred <- states_map |>
  left_join(win_pred, by = "region")

fig1 <- ec24_pred |>
  ggplot(aes(long, lat, group = group)) +
  geom_polygon(aes(fill = winner), color = "#000033") +
  scale_fill_manual(values = c("#7CB9E8", "#FF6680")) +
  ggtitle("Fig. VII - Electoral College Prediction, 2024") +
  my_map_theme() +
  theme(strip.text = element_text(size = 12), 
        aspect.ratio = 0.75)

# Re-run for 2020

mod_lm_dem20 <- lm(state_pv2p ~ ics + prev_admin + poll_week_1 + state_pv2p_lag + mean_support, data = d_train20)
summary(mod_lm_dem20)

mod_lm_dem_simp20 <- lm(state_pv2p ~ ics + prev_admin + poll_week_1 + state_pv2p_lag + mean_support, data = d_train20)
summary(mod_lm_dem_simp20) # adjusted r-squared = 0.7683, all vars except poll significant

d_test_simp20 <- d_test20 |>
  select(state, year, all_of(simp.vars))

simp_pred_dem20 <- predict(mod_lm_dem_simp20, d_test_simp20)
simp_pred_dem_interval20 <- predict(mod_lm_dem_simp20, d_test_simp20, interval = "prediction", level = 0.8)
simp_pred_dem_interval20 <- as.data.frame(simp_pred_dem_interval20)
s

win_pred20 <- data.frame(state = d_test_simp20$state,
                       year = rep(2020, length(d_test_simp20$state)),
                       simp_pred_dem = simp_pred_dem20,
                       winner = ifelse(simp_pred_dem20 > 50, "Democrat", "Republican")) |>
  left_join(d_ec, by = c("state", "year")) |>
  mutate(
    region = tolower(state)
  )

win_pred20 <- win_pred20 |> 
  mutate(
    biden_win = ifelse(winner == "Democrat", 1, 0),
    biden_ec = (biden_win*electors)
  )
sum(win_pred20$biden_ec) # 409 Biden EC (not counting DC)

win_pred20 <- win_pred20 |> cbind(lower = simp_pred_dem_interval20$lwr, upper = simp_pred_dem_interval20$upr)

ec20_pred <- states_map |>
  left_join(win_pred20, by = "region")

figx <- ec20_pred |>
  ggplot(aes(long, lat, group = group)) +
  geom_polygon(aes(fill = winner), color = "#000033") +
  scale_fill_manual(values = c("#7CB9E8", "#FF6680")) +
  ggtitle("Fig. III - Electoral College Prediction, 2020") +
  my_map_theme() +
  theme(strip.text = element_text(size = 12), 
        aspect.ratio = 0.75)

ec20_rescaled <- ec20_pred |>
  mutate(
    simp_pred_dem = simp_pred_dem - 50
  )
figy <- ec20_rescaled |>
  ggplot(aes(long, lat, group = group)) +
  geom_polygon(aes(fill = simp_pred_dem), color = "#000033") +
  scale_fill_gradient2(high = "#7CB9E8", # high (positive) values represent a greater average Republican vote share
                       low = "#FF6680", # low (negative) values represent a greater average Democratic vote share
                       mid = "white", # competitive margins approach zero and display as white / light-colored
                       name = "margin",
                       breaks = c(-50, -25, 0, 25, 50),
                       limits = c(-25, 25)) +
  ggtitle("Fig. IV - State-Level Vote Margin, 2020") +
  my_map_theme() +
  theme(strip.text = element_text(size = 12), 
        aspect.ratio = 0.75)

# Re-run for 2016

mod_lm_dem16 <- lm(state_pv2p ~ ics + prev_admin + poll_week_1 + state_pv2p_lag + mean_support, data = d_train16)
summary(mod_lm_dem16)

mod_lm_dem_simp16 <- lm(state_pv2p ~ ics + prev_admin + poll_week_1 + state_pv2p_lag + mean_support, data = d_train16)
summary(mod_lm_dem_simp16) # adjusted r-squared = 0.7683, all vars except poll significant

d_test_simp16 <- d_test16 |>
  select(state, year, all_of(simp.vars))

simp_pred_dem16 <- predict(mod_lm_dem_simp16, d_test_simp16)
simp_pred_dem_interval16 <- predict(mod_lm_dem_simp16, d_test_simp16, interval = "prediction", level = 0.8)
simp_pred_dem_interval16 <- as.data.frame(simp_pred_dem_interval16)
s

win_pred16 <- data.frame(state = d_test_simp16$state,
                       year = rep(2016, length(d_test_simp16$state)),
                       simp_pred_dem = simp_pred_dem16,
                       winner = ifelse(simp_pred_dem16 > 50, "Democrat", "Republican")) |>
  left_join(d_ec, by = c("state", "year")) |>
  mutate(
    region = tolower(state)
  )

win_pred16 <- win_pred16 |> 
  mutate(
    clinton_win = ifelse(winner == "Democrat", 1, 0),
    clinton_ec = (clinton_win*electors)
  )
sum(win_pred16$clinton_ec) # 243 Clinton EC (not counting DC)

win_pred16 <- win_pred16 |> cbind(lower = simp_pred_dem_interval16$lwr, upper = simp_pred_dem_interval16$upr)

ec16_pred <- states_map |>
  left_join(win_pred16, by = "region")

figza <- ec16_pred |>
  ggplot(aes(long, lat, group = group)) +
  geom_polygon(aes(fill = winner), color = "#000033") +
  scale_fill_manual(values = c("#7CB9E8", "#FF6680")) +
  ggtitle("Fig. V - Electoral College Prediction, 2016") +
  my_map_theme() +
  theme(strip.text = element_text(size = 12), 
        aspect.ratio = 0.75)

ec16_rescaled <- ec16_pred |>
  mutate(
    simp_pred_dem = simp_pred_dem - 50
  )
figzb <- ec16_rescaled |>
  ggplot(aes(long, lat, group = group)) +
  geom_polygon(aes(fill = simp_pred_dem), color = "#000033") +
  scale_fill_gradient2(high = "#7CB9E8", # high (positive) values represent a greater average Republican vote share
                       low = "#FF6680", # low (negative) values represent a greater average Democratic vote share
                       mid = "white", # competitive margins approach zero and display as white / light-colored
                       name = "margin",
                       breaks = c(-50, -25, 0, 25, 50),
                       limits = c(-25, 25)) +
  ggtitle("Fig. VI - State-Level Vote Margin, 2016") +
  my_map_theme() +
  theme(strip.text = element_text(size = 12), 
        aspect.ratio = 0.75)
```

```{r fig a code, echo=FALSE, include=FALSE, warning=FALSE}
# sequester training data
pop_pre <- d_pop_tfmc |> filter(year <= 2020)

# run cross-validation 1k times
set.seed(19709)
pop_out_samp_errors <- sapply(1:1000, function(i) {
  years_out_samp <- sample(pop_pre$year, 6)
  mod <- lm(pv2p ~ ics + prev_admin + poll_week_1 + pv2p_lag,
            pop_pre[!(pop_pre$year %in% years_out_samp),])
  out_samp_pred <- predict(mod, pop_pre[pop_pre$year %in% years_out_samp,])
  out_samp_truth <- pop_pre$pv2p[pop_pre$year %in% years_out_samp]
  mean(out_samp_pred - out_samp_truth)
})

# mean of out-of-sample residuals
mean(abs(pop_out_samp_errors)) # 2.615

# histogram of prediction errors (out-of-sample residuals)
figa <- ggplot() +
  geom_histogram(mapping = aes(x = pop_out_samp_errors), fill = "#9999CC", binwidth = 2) +
  xlab("Out-of-Sample Residuals") +
  ylab("Count") +
  ggtitle("Fig. II - Out-of-Sample Error (Popular Vote)") +
  my_theme()
```

```{r table 2 code, echo=FALSE, warning=FALSE, include=FALSE}
swing_pred <- win_pred |>
  filter(state == "Arizona" | state == "Georgia" | state == "Michigan" | state == "Pennsylvania" | state == "Nevada" | state == "North Carolina" | state == "Wisconsin") |>
  select(state, simp_pred_dem, lower, upper, winner) |>
  rename(`Dem Two-Party Vote Share` = simp_pred_dem, Lower = lower, Upper = upper, State = state, Winner = winner)
table2 <- knitr::kable(swing_pred, digits = 2)
```

```{r polls table code, echo=FALSE, include=FALSE, warning=FALSE}
poll_comp <- tibble(State = c("Indiana", "Massachusetts", "Nebraska", "Washington"),
       `Dem Two-Party Vote Share (Polls Available)` = c("41.47", "65.48", "40.19", "59.28"),
       `Dem Two-Party Vote Share (Polls Imputed)` = c("41.76", "66.95", "40.18", "59.79"))

table_poll <- knitr::kable(poll_comp)
```

```{r fig 2 code, echo=FALSE, include=FALSE, warning=FALSE}
ec24_rescaled <- ec24_pred |>
  mutate(
    simp_pred_dem = simp_pred_dem - 50
  )
fig2 <- ec24_rescaled |>
  ggplot(aes(long, lat, group = group)) +
  geom_polygon(aes(fill = simp_pred_dem), color = "#000033") +
  scale_fill_gradient2(high = "#7CB9E8", # high (positive) values represent a greater average Republican vote share
                       low = "#FF6680", # low (negative) values represent a greater average Democratic vote share
                       mid = "white", # competitive margins approach zero and display as white / light-colored
                       name = "margin",
                       breaks = c(-50, -25, 0, 25, 50),
                       limits = c(-25, 25)) +
  ggtitle("Fig. VIII - State-Level Vote Margin, 2024") +
  my_map_theme() +
  theme(strip.text = element_text(size = 12), 
        aspect.ratio = 0.75)
```

# Introduction

Welcome to my final prediction for the 2024 presidential election!

With early voting well underway and only a few days left until Election Day, the race remains stubbornly in the toss-up category, according to expert forecasters such as the [Cook Political Report](https://www.cookpolitical.com/ratings/presidential-race-ratings) and [Sabatos Crystal Ball](https://centerforpolitics.org/crystalball/2024-president/). My prediction, unfortunately, is no exception to this trend. While I will predict a winner in todays post, the predictive intervals produced by both my popular vote and electoral college models encompass scenarios in which either candidate could win.

# My 2024 Model

My model uses a simple OLS regression, with the following explanatory variables: a dummy indicator of whether the candidate served in a previous presidential administration, Q2 consumer sentiment, lagged vote share, and a recent polling average. The outcome variable is the two-party vote share for the Democratic candidate.

My model is inspired by Abramowitzs Time For Change model, though I changed the specific indicators (incumbency to previous administration, GDP to ICS, and June approval ratings to recent polls) and added lagged vote share.

Here is the formula for the popular vote regression:

$Y = \beta_0 + \beta_1A + \beta_2C + \beta_3L + \beta_4P + \epsilon$

And here is the formula for the electoral college regression:

$Y = \beta_0 + \beta_1A + \beta_2C + \beta_3V + \beta_4P + \beta_5S + \epsilon$

I will address each of the variables and their coefficients in the following sections, but first I should briefly discuss adjusted R-squared, a measure of the models overall performance. Adjusted R-squared describes the percent of the variation in Democratic two-party vote share that can be explained by the independent variables. For the popular vote, this is about 71%, while the electoral college model, probably since it is more finely tuned to state-level conditions, explains about 78%. While there are ways to inflate R-squared and adjusted R-squared, I think this 71%-78% range, even though it does not account for all possible factors, is respectable for the purposes of this prediction. After all, a sky-high R-squared might seem desirable, but it could indicate overfitting, or hyper-sensitivity to trends in the historical data, which could impair the predictive power of the model.

```{r reg_table code, echo=FALSE, include=FALSE, warning=FALSE}
regression <- data.frame(Statistic = c("Intercept", "Index of Consumer Sentiment (C)", "National Poll Average, 1 Week Left (P)", "Candidate Was In Previous Administration (A)", "Previous Year National Two-Party Vote Share (L)", "Previous Year State Two-Party Vote Share (V)", "State Poll Average or Imputed Support, All-Time (S)", "Number of Observations", "R-squared", "Adjusted R-squared"),
                        `Popular Vote` = c(15.806, 0.008, 0.674, -1.358, 0.074, "not included", "not included", 13, 0.806, 0.709),
                        `Electoral College` = c(8.105, -0.035, -0.033, -4.307, "not included", 0.600, 0.425, 559, 0.779, 0.777))

renamed_cols <- c(" ", "Popular Vote Model", "Electoral College Model")

reg_table <- knitr::kable(regression, col.names = renamed_cols)
```

```{r print reg_table, echo=FALSE, include=TRUE, warning=FALSE}
reg_table
```

## Previous Administration

I have already written at length about the problem of incumbency in this years race, but this, at least, bears repeating: the boons and burdens of incumbency are split between Trump and Harris. He has impeccable name recognition, has been the leader of his party for eight years, and hopes to capitalize off of nostalgia for his first term in office. She inherited Bidens extensive campaign machinery and must defend the current administrations record, including on unflattering issues like inflation, immigration, and war in the Middle East. Since my model only predicts Democratic vote share, Trumps incumbency status is left an open question. Harris, however, must be directly addressed.

I have chosen to use a previous administration dummy because I think it comes closest to describing Harris as a non-incumbent who is nevertheless closely linked to the sitting president. A binary indicator for incumbent party membership could also be used here, but it risks underplaying the extent to which voters will evaluate Harris as Bidens second-in-command. Statistical tests are of little use in this instance, since Bidens last-minute removal from the race is anomalous. Whether you lend any credibility to this model, then, depends on how you take this judgment call.

Both coefficients  for the popular vote and electoral college models  are negative when the previous administration variable is 1. According to the models, then, being a member of a previous presidential administration is associated with decreased vote share. However, this relationship is only significant in the electoral college model, where members of previous administrations are expected to receive about 4.3 fewer points compared to other candidates.

## Index of Consumer Sentiment

My economic indicator of choice is another slight twist on the Time For Change formula. Instead of GDP, I opted for a measure of consumer sentiment, or respondents professed impressions of the state of the economy, with the hope of more directly assessing the judgments underlying economic retrospective voting. The choice of this indicator assumes a model of voter behavior in which voters do weigh their personal impressions of the economy but are not substantially subconsciously influenced by fluctuations in GDP that may be influencing economic conditions behind the scenes.

```{r print ics_outliers, echo=FALSE, include=TRUE, warning=FALSE}
ics_outliers
```

The same story we observed for the previous administration variable repeats itself with consumer sentiment. In the popular vote model, the coefficient is substantively low and insignificant. In the electoral college model, it is significant at the 0.05 level and has a small negative estimate of -0.0035. Counterintuitively, this coefficient implies that Democratic vote share drops as Americans outlook on the economy improves. If this trend does hold true in reality, it could result from an increased desire for welfare and stimulus programs (associated with the Democratic party) during times of economic hardship and a desire for more limited government intervention (associated with the Republican party) when the economy seems strong.

## Lagged Vote Share

Continuity over time is the very basic principle behind the inclusion of a lagged vote share variable. Vavrecks signature argument that the United States is experiencing a process of calcification (from our GOV 1347 class discussion), whereby partisans are becoming entrenched in their views and more homogeneous among themselves, would suggest that previous cycle vote share will only become a better predictor of voter behavior as the electorate calcifies.

As seems to be the trend for most of these variables, the national lagged vote share variable is insignificant and substantively small, while the state-level lagged vote share is highly significant and has an estimate of 0.6, meaning every additional point for the Democratic party in the preceding election is expected to correspond to 0.6 additional points in the next race. This would be moderately good news for the Harris campaign, since it suggests that she would inherit a slight head start in those battlegrounds which Biden won in 2020.

## Poll Average

With polling, we come to the first major distinction between national- and state-level models. The most recent national polling average from FiveThirtyEights aggregation  currently set at one week before Election Day, though the polls themselves may have been administered earlier  is featured in both models. I decided to limit national polls to the most recent weekly average in recognition of [Gelman and King](https://doi.org/10.1017/S0007123400006682)'s claim that polls become more reliable closer to Election Day, as voters converge on their underlying preferences. However, I also chose to take an all-time average of state-level polls for my electoral college. While ideally I would like to have plentiful recent polls from all states, state-level polls are fewer in number than national polls, and I therefore opted not to restrict my sample further by removing old polls. Hopefully, this two-pronged approach gives my electoral college model the best of both worlds: the recent national average should act as a temperature check for Harriss overall trajectory across the country and the long-term state average should offer an insight into state-level preferences that is more resistant to short-term fluctuations.

Unfortunately, in some cases, state-level polling is not yet available. For these states, I imputed a poll average by applying their deviation from the national popular vote in 2020 to the current national average. In this past week, polling data became available for four states that I had previously imputed, so I took the opportunity to compare the results of my imputation to the actual polls.

```{r print table_poll, echo=FALSE, include=TRUE, warning=FALSE}
table_poll
```

As the above table indicates, my imputation procedure generated very similar predictions to what my model is now predicting given new polling data. While none of these states are highly contested  and, for that matter, none of remaining imputed states are battlegrounds  it is reassuring that the biggest shift was just 1.47 points in Massachusetts.

In the popular vote model, the national poll average (P) just barely misses the threshold for significance at the 0.05 level. Its coefficient, 0.674, suggests that a one-point increase in Democratic support in the national poll average corresponds with a 0.674 increase in Democratic vote share. In the electoral college model, the national average has a small negative coefficient that is not significant. Interestingly, the coefficient is negative, which, if true, would imply that Democratic vote share in the states slightly declines as support in the national polls rises. However, the state-level average is highly significant and has a coefficient of 0.425, which could be a substantial pay-off in state-level vote share.

# Out-of-Sample Tests

To further assess the model, I conducted a cross-validation procedure on the popular vote model and re-ran the electoral college model to predict the 2016 and 2020 election cycles.

Figure II, below, summarizes the out-of-sample residuals from the cross-validation exercise. The distribution is roughly normal, as expected, and the mean of the residuals is about 2.615. In the context of a bitterly close election, of course, 2.615 points is nothing to sniff at. However, given the range in observed vote share found in the historical data is around 16.6 points, and that this degree of out-of-sample error is in line with the predictive models I have produced in previous weeks, I have continued with this model.

```{r print figa, echo=FALSE, include=TRUE, warning=FALSE}
figa
```

Next, I created an electoral college predictions for 2020. Figure III, below, illustrates which states my model would expect Biden to win and lose in 2020.

```{r print figx, echo=FALSE, include=TRUE, warning=FALSE}
figx
```

There are, as I am sure you have noticed, several things wrong with this map. Not only does it predict a blowout victory for Biden with 409 electoral votes, but it does so by handing him improbable wins in states like Texas and Florida. On the other hand, Figure IV, which displays Bidens predicted margins in each state, makes these ludicrous upfront results seem a bit more sensible.

```{r print figy, echo=FALSE, include=TRUE, warning=FALSE}
figy
```

The margins map, above, at least gives margin a very narrow margin of victory in Texas and Florida. Rather than throwing all of the states randomly out of whack, the model seems to have preserved the basic relative values of the states, but shifted the whole prediction slightly leftward. This is particularly interesting since, as we will soon see, my 2024 prediction from the same model appears to have the opposite skew.

As a final check, I used my model to predict the 2016 election, with similarly bizarre results. While Figure V, below, shows that my model did predict a Trump victory, it shortchanged Clinton in Colorado, New Mexico, and Virginia, despite handing her wins in the Blue Wall states.

```{r print figza, echo=FALSE, include=TRUE, warning=FALSE}
figza
```

Once again, Figure VI indicates that these mistakes come from states with very narrow predictive margins.

```{r print figzb, echo=FALSE, include=TRUE, warning=FALSE}
figzb
```

Ultimately, my model is only as good as the data that it receives. These alternate-reality maps of 2016 and 2020 are a reminder that my model rests on the  possibly fragile  hope that the same pollsters who veered in the wrong direction, say by [underestimating Trump in 2016](https://www.pewresearch.org/short-reads/2016/11/09/why-2016-election-polls-missed-their-mark/), will not make a similar systematic error in 2024.

# Popular Vote Results

```{r print table1, echo=FALSE, include=TRUE, warning=FALSE}
table1
```

# Electoral College Results

```{r print fig1, echo=FALSE, include=TRUE, warning=FALSE}
fig1
```

```{r print table3, echo=FALSE, include=TRUE, warning=FALSE}
table3
```

```{r print table2, echo=FALSE, include=TRUE, warning=FALSE}
table2
```

```{r print fig2, echo=FALSE, include=TRUE, warning=FALSE}
fig2
```

