---
title: "Campaign Narrative"
author: Ella Trembanis
date: '2024-12-07'
output:
  html_document:
    df_print: paged
categories: []
tags: []
slug: "campaign-narrative"
---

```{r load libraries, echo=FALSE, include=FALSE, warning=FALSE}
library(censable)
library(geofacet)
library(ggpubr)
library(ggthemes)
library(haven)
library(kableExtra)
library(maps)
library(mgcv)
library(mgcViz)
library(RColorBrewer)
library(readstata13)
library(scales)
library(sf)
library(spData)
library(stargazer)
library(tidygeocoder)
library(tidyverse)
library(tigris)
library(tmap)
library(tmaptools)
library(viridis)
```

```{r custom theme, echo=FALSE, include=FALSE, warning=FALSE}
my_theme <- function() {
  theme(
  # set panel features
  panel.border = element_rect(color = "#000033", fill = NA, linetype = 1),
  panel.background = element_rect(fill = "#FFFFFF"),
  panel.grid.major.x = element_line(color = "#9999CC", linetype = 2, size = 0.2),
  panel.grid.minor.x = element_blank(),
  panel.grid.major.y = element_line(color = "#9999CC", linetype = 2, size = 0.2),
  panel.grid.minor.y = element_blank(),
  # set text and axis features
  text = element_text(color = "#000033", family = "Courier"),
  title = element_text(color = "#000033", face = "bold", family = "Courier"),
  axis.ticks = element_line(color = "#000033"),
  # set legend features
  legend.position = "bottom",
  legend.background = element_rect(color = "#000033", fill = NA, linetype = 1)
  )
}
```

```{r simplified custom theme, echo=FALSE, include=FALSE, warning=FALSE}
my_map_theme <- function() {
  theme(
    # set panel features
    panel.background = element_rect(fill = "#FFFFFF"),
    # set text features
    text = element_text(color = "#000033", family = "Courier"),
    title = element_text(color = "#000033", face = "bold", family = "Courier"),
    # set legend features
    legend.position = "bottom",
    legend.title.position = "top",
    legend.background = element_rect(color = "#000033", fill = NA, linetype = 1),
    # set axis features to not display
    axis.title = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank()
  )
}
```

```{r read data, echo=FALSE, include=FALSE, warning=FALSE}
# Read 2024 results datasets. 
d_state_2024 <- read_csv("/Users/ellatrembanis/Desktop/Gov 1347/election-blog/election-blog/Week10/data/state_votes_pres_2024.csv")[-1, 1:6]
d_county_2024 <- read_csv("/Users/ellatrembanis/Desktop/Gov 1347/election-blog/election-blog/Week10/data/county_votes_pres_2024.csv")[-1, 1:6]
d_county_2020 <- read_csv("/Users/ellatrembanis/Desktop/Gov 1347/election-blog/election-blog/Week10/data/county_votes_pres_2020.csv")[-1, 1:6]

# Process 2024 state and county-level data. 
d_state_2024 <- d_state_2024 |> 
  mutate(FIPS = as.numeric(FIPS), 
         votes_trump = as.numeric(`Donald J. Trump`), 
         votes_harris = as.numeric(`Kamala D. Harris`), 
         votes = as.numeric(`Total Vote`), 
         trump_pv = votes_trump/votes, 
         harris_pv = votes_harris/votes, 
         trump_2pv = votes_trump/(votes_trump + votes_harris), 
         harris_2pv = votes_harris/(votes_trump + votes_harris)) |> 
  mutate(winner = case_when(votes_trump > votes_harris ~ "REP", 
                            .default = "DEM")) |> 
  select(FIPS, `Geographic Name`, `Geographic Subtype`, votes_trump, votes_harris, votes, 
         winner, trump_pv, harris_pv, trump_2pv, harris_2pv)

d_county_2024 <- d_county_2024 |>
  mutate(FIPS = as.numeric(FIPS),
         votes_trump = as.numeric(`Donald J. Trump`), 
         votes_harris = as.numeric(`Kamala D. Harris`), 
         votes = as.numeric(`Total Vote`), 
         trump_pv = votes_trump/votes, 
         harris_pv = votes_harris/votes, 
         trump_2pv = votes_trump/(votes_trump + votes_harris), 
         harris_2pv = votes_harris/(votes_trump + votes_harris)) |> 
  mutate(winner = case_when(votes_trump > votes_harris ~ "REP", 
                            .default = "DEM")) |> 
  select(FIPS, `Geographic Name`, `Geographic Subtype`, votes_trump, votes_harris, votes, 
         winner, trump_pv, harris_pv, trump_2pv, harris_2pv)

d_county_2020 <- d_county_2020 |> 
  mutate(FIPS = as.numeric(FIPS),
         votes_trump_2020 = as.numeric(`Donald J. Trump`), 
         votes_biden_2020 = as.numeric(`Joseph R. Biden Jr.`), 
         votes_2020 = as.numeric(`Total Vote`), 
         trump_pv_2020 = votes_trump_2020/votes_2020, 
         biden_pv_2020 = votes_biden_2020/votes_2020, 
         trump_2pv_2020 = votes_trump_2020/(votes_trump_2020 + votes_biden_2020), 
         biden_2pv_2020 = votes_biden_2020/(votes_trump_2020 + votes_biden_2020)) |> 
  mutate(winner_2020 = case_when(votes_trump_2020 > votes_biden_2020 ~ "REP", 
                            .default = "DEM")) |> 
  select(FIPS, `Geographic Name`, `Geographic Subtype`, votes_trump_2020, votes_biden_2020, votes_2020, 
         winner_2020, trump_pv_2020, biden_pv_2020, trump_2pv_2020, biden_2pv_2020)
```

```{r map set-up, echo=FALSE, include=FALSE, warning=FALSE}
# Sequester state and county-level map.
states_2024 <- states(cb = TRUE, year = 2023) |> 
  shift_geometry() |> 
  mutate(GEOID = as.numeric(GEOID)) |> 
  left_join(d_state_2024, by = c("GEOID" = "FIPS")) |> 
  drop_na()
counties_2024 <- counties(cb = TRUE, resolution = "5m", year = 2023) |> 
  shift_geometry() |> 
  mutate(GEOID = as.numeric(GEOID)) |> 
  left_join(d_county_2024, by = c("GEOID" = "FIPS")) |> 
  left_join(d_county_2020, by = c("GEOID" = "FIPS")) |>
  mutate(shift = (trump_pv - trump_pv_2020) * 100, 
         `shift direction` = case_when(shift > 0 ~ "REP", 
                               shift < 0 ~ "DEM", 
                               TRUE ~ "No Change"),
         centroid = st_centroid(geometry), 
         centroid_long = st_coordinates(centroid)[,1],
         centroid_lat = st_coordinates(centroid)[,2],
         scale_factor = 1e4, 
         end_long = centroid_long + scale_factor * shift,
         end_lat = centroid_lat + scale_factor * shift) |>
  drop_na()
county_pop_2024 <- read_csv("/Users/ellatrembanis/Desktop/Gov 1347/election-blog/election-blog/Week10/data/PopulationEstimates.csv") |> 
  mutate(FIPStxt = as.numeric(FIPStxt)) |>
  select(FIPStxt, POP_ESTIMATE_2023)
counties_2024 <- counties_2024 |> 
  left_join(county_pop_2024, by = c("GEOID" = "FIPStxt"))
```

```{r ne table code, echo=FALSE, include=FALSE, warning=FALSE}
ne_results <- tibble(`Unit of Analysis` = c("Statewide", "CD-1", "CD-2", "CD-3"),
                     "Harris Two-Party Vote Share (2024)" = c(39.4, 43.4, 52.3, 22.8),
                     "My Prediction" = c("40.2", "NA", "NA", "NA"),
                     "538's Prediction" = c(41.3, 41.6, 53.5, 22.6),
                     "CPR Rating" = c("Solid R", "Solid R", "Likely D", "Solid R"),
                     "Biden Two-Party Vote Share (2020)" = c(39.2,43.3, 52.2, 23.1))

ne_table <- knitr::kable(ne_results)
```

```{r ne map code, echo=FALSE, include=FALSE, warning=FALSE}
# Check county-level shifts in Nebraska between 2024 and 2020. 
ne_map <- counties_2024 |> 
  filter(STATE_NAME == "Nebraska") |> 
  ggplot() +
  geom_sf(fill = "#FCFCFC", color = "#000033") +  # Base map
  geom_label(aes(x = centroid_long, y = centroid_lat-1e4, label = NAME),
            size = 2,  # Adjust size as needed
            color = "black", hjust = 0.5, vjust = -0.005) + 
  geom_curve(aes(x = centroid_long, 
                 y = centroid_lat,
                 xend = end_long, 
                 yend = end_lat,
                 color = `shift direction`),
             arrow = arrow(length = unit(0.1, "cm"), type = "closed"),  # Smaller arrowhead
             curvature = 0.2,  # Add a slight curve to each arrow
             size = 0.3) +
  scale_color_manual(values = c("DEM" = "#7CB9E8", "REP" = "#FF6680")) +
  my_map_theme() +
  labs(title = "Presidential Voting Shifts by County (2020-2024)")
```

# Campaigning in the Cornhusker State

Welcome to the very final installment of my 2024 election blog!

Today, I will be zeroing in on Nebraska and dissecting the campaign activity in the state in the hopes of figuring out why forecasters (including myself!) might have missed the true outcome.

For starters, the map below depicts the county-level swing from 2020. Almost all of Nebraska's counties experienced an increase in Republican vote share relative to the 2020 cycle.

The few exceptions -- [Kimball](https://irp.cdn-website.com/82f946ac/files/uploaded/General_Election_Results-Pres_2024-c4b5ed95.pdf), [Boyd](https://countyelectionresults.nebraska.gov/election_files/Boyd/2024_General_official_results.pdf), [Brown](https://electionresults.nebraska.gov/resultsCTY.aspx?type=PC&rid=12465&osn=93&map=CTY), [Sheridan](https://electionresults.nebraska.gov/resultsCTY.aspx?type=PC&rid=12465&osn=93&map=CTY) and [Deuel](https://countyelectionresults.nebraska.gov/election_files/Deuel/2024_General_official_results.pdf) -- are all ruby red, sparsely populated rural counties in Nebraska's third district. Their slight leftward swings could just be random noise that their electorates are simply too small to drown out.

On the other hand, the red swing is present but modest in Omaha's Douglas County and the surrounding areas which make up Nebraska's second and first congressional districts. This is, of course, despite the fact that Harris did win NE-2's one electoral vote. While it is impossible to know exactly what the magnitude of the swing might have been if either of the campaigns pursued different strategies in NE-2, it is a question worth keeping in mind -- especially since, as we shall see, both Trump and Harris had their eyes on Nebraska's famous "blue dot."

```{r print ne map, echo=FALSE, include=TRUE, warning=FALSE}
ne_map
```

Before we get any further, though, I should establish how the predictions stacked up against this reality. Since my model was reliant on historical polling data, and I could not source sufficient observations for Nebraska’s congressional districts, I only generated a statewide prediction. To compensate for this missingness, I have also included predictions from [538](https://projects.fivethirtyeight.com/2024-election-forecast/) and ratings from the [Cook Political Report](https://www.cookpolitical.com/ratings/presidential-race-ratings).

```{r print ne table, echo=FALSE, include=TRUE, warning=FALSE}
ne_table
```

Ultimately, all of the top-line outcomes were predicted correctly: Trump won statewide and in the first and third congressional districts, and Harris won the second congressional district. My statewide prediction was slightly closer to the outcome than 538’s, perhaps due to overfitting in their higher-dimensional model. In any case, both my model and the 538 model slightly overestimated Harris’s statewide vote share.

If we break the 538 predictions down by congressional district, we get what appear to be three quite different stories: the model was almost dead-on in NE-3, overestimated Harris by 1.2 points in NE-2, and underestimated her by 1.8 points in NE-1. Since 538’s model performed remarkably well in NE-3, a district which also received very little attention from either presidential campaign, I will confine my remarks to the other two districts and the statewide result, with special attention to NE-2.

# Nebraska: An Overview

## Demographics, Geography, & Electoral History

Nebraska is a [Republican bastion](https://www.270towin.com/states/Nebraska) in the heart of the Great Plains. A Democratic presidential candidate has not won statewide in Nebraska since Lyndon B. Johnson in 1964 and the state has consistently elected Republican governors since 1998.

Nebraska does tick some boxes in the “Republican stereotypes” column: it is highly rural, [ranking 43rd](https://www.census.gov/quickfacts/geo/chart/NE/PST045223) in the nation for population density, and has a large proportion of non-Hispanic whites ([approximately 76%](https://www.census.gov/quickfacts/fact/table/NE/PST045223)). Nonetheless, the state is firmly [in the middle of the pack](https://www.census.gov/quickfacts/geo/chart/NE/PST045223) on other demographic indicators, including household income, educational attainment, and computer/broadband access.

## Red State, Purple Dot

No doubt the most famous feature of Nebraska’s electoral system is its method of splitting its five electoral votes. Unlike most states, which award all of their votes to whichever candidate receives a plurality in the state-level popular vote, Nebraska awards just two votes to the statewide winner and one vote to the winner of each of its three congressional districts.

Compared to winner-take-all, this configuration is a boon for Democrats; in 2008 and 2020, Nebraska’s second congressional district awarded one vote to the Democratic candidate. However, the idea that NE-2 is a reliable “blue dot” in the ruby-red state is at least a little overstated. This is a district, after all, that voted for a Republican presidential candidate as recently as [2016](https://nebraskaexaminer.com/2024/08/10/nebraskas-2nd-district-steps-back-into-presidential-spotlight-after-crazy-month/) – a district that, to this day, is represented in the House by a moderate Republican. Even in [Omaha](https://www.politico.com/news/magazine/2024/10/14/nebraska-2nd-district-battleground-00183400), the city nominally driving NE-2’s “blue dot” status, the mayor is a Republican and registered Republicans outnumber registered Democrats. As the national electorate calcifies, and “red” and “blue” states become steadily more predictable, the so-called “blue dot” has more in common with the kind of Midwestern battleground that [Hopkins](https://doi.org/10.1017/9781108123594.006) described in the wake of the 2016 election: purple, elastic, and torn between urban and rural constituents.

Earlier this fall, Nebraska Republicans, with Trump’s blessing, [launched](https://nebraskaexaminer.com/2024/08/10/nebraskas-2nd-district-steps-back-into-presidential-spotlight-after-crazy-month/) a campaign to re-institute winner-take-all. The push [fell short](https://nebraskaexaminer.com/2024/09/23/state-sen-mike-mcdonnell-deflates-gop-hopes-for-nebraska-winner-take-all-in-2024/) in the state legislature thanks to a pivotal “no” vote from Republican Mike McDonnell, a former Omaha Democrat. McDonnell, however, did not set aside the possibility of changing Nebraska’s electoral vote allocation method in time for the next presidential election.

## Elsewhere on the Ballot

Of course, the presidential campaign did not happen in isolation. Nebraska voters faced other important down-ballot choices on Election Day, and the state generally affirmed its preference for Republican officeholders and conservative policies.

Republican Senator Deb Fischer was [re-elected](https://ballotpedia.org/United_States_Senate_election_in_Nebraska,_2024) over Independent challenger Dan Osborn. Osborn’s middle-of-the-road campaign, which emphasized his nonpartisanship and his working-class roots, made a reasonably competitive race out of what might otherwise have been a slam dunk for Fischer. There are difficulties in drawing analogies between this race and the presidential contest, since both candidates [attempted](https://www.washingtonexaminer.com/news/campaigns/congressional/3214917/dan-osborn-nebraska-senate-race-independent/) to align themselves with Trump – though only Fischer received his coveted endorsement.

Nebraskans also weighed in on a range of ballot initiatives, and the results are a fascinating mixed bag. Bucking the post-Dobbs [trend](https://time.com/7173410/abortion-ballot-results-2024-election/) of enthusiasm for reproductive rights on the ballot, Nebraskans [struck down](https://ballotpedia.org/Nebraska_elections,_2024) a measure to enshrine a right to abortion in the state constitution and supported a ban for abortions after the first trimester of pregnancy. Interestingly, the margin was substantially lower for the former initiative (3 points versus 10 points) and fewer voters responded to the latter question. Perhaps it was easier to mobilize voters around the relatively clear terms of the ban – which specifies the trimester of concern as well as exceptions for rape, incest, and the health of the mother – compared to the somewhat more vague language of “fetal viability” used in the proposed constitutional amendment.

Nevertheless, voters also opted to require paid sick leave, to legalize medical marijuana (and establish a regulatory body to oversee its usage), and to overturn a law which would provide vouchers to qualified private school students.

Finally, voters in the state’s all-important second Congressional District [chose](https://ballotpedia.org/Nebraska%27s_2nd_Congressional_District_election,_2024) to return Republican incumbent Don Bacon to the House of Representatives. As a testament to the purple-ness of the district, however, Bacon’s opponent, Democrat Tony Vargas, lost by a slim 1.8-point margin.

# 2024 Campaign Activity

From the beginning, it was clear that NE-2 was a second-rate priority for both the Trump and Harris campaigns. True, there were the inevitable [flutterings](https://www.usatoday.com/story/news/politics/elections/2024/09/12/2024-presidential-race-nebraska-blue-dot/75018361007/) about how the polls were just so close that it might come down to NE-2's single vote and national Republicans mustered some strength in their winner-take-all bid, but that energy quickly fizzled.

Although Nebraska spending estimates for the Trump campaign [have not yet been released](https://www.fec.gov/data/spending-bythenumbers/) by the Federal Election Commission, Harris [allegedly outspent](https://www.ketv.com/article/election-week-the-presidential-ground-game-in-omaha/62798017) Trump in NE-2, [employed more staff members](https://www.ketv.com/article/election-week-the-presidential-ground-game-in-omaha/62798017), and [ran the lion’s share of presidential campaign ads](https://mediaproject.wesleyan.edu/releases-103124/) in the Omaha DMA. Neither candidate held their own rallies in the Cornhusker state, opting instead to [send surrogates](https://abcnews.go.com/Politics/blue-dot-nebraska-draws-boldface-political-names/story?id=114967525), including Tim and Gwen Walz, Doug Emhoff, Robert F. Kennedy Jr., and Tulsi Gabbard. Both campaigns had active door-knocking and phone-banking campaigns, with Trump Force 47 – Nebraska’s joint office for the RNC and the Trump campaign – apparently pouring most of its efforts into [mobilizing](https://www.politico.com/news/magazine/2024/10/14/nebraska-2nd-district-battleground-00183400) extremely low-turnout voters.

# Conclusions

## Redistricting

My first theory has less to do with voter or candidate behavior, but since it has the potential to explain the discrepancy between 538’s NE-1 and NE-2 prediction errors, I think it is worth mentioning here.

After the release of the 2020 Census, Nebraska’s Republican-run government [redrew](https://ballotpedia.org/Redistricting_in_Nebraska_after_the_2020_census) the boundaries of its three congressional districts. The changes were for the most part subtle, but they involved extending NE-2 further west, into the state’s more rural counties. This change presents a unique challenge for forecasters, whose task mostly involves predicting outcomes in states with long-fixed borders. While 2024 polling in NE-2 should in theory incorporate these newly eligible voters’ intentions (though of course this could be complicated if rural voters prove harder to contact than their urban and suburban counterparts), any historical trends that the model relies on will not be from exactly the same place. Moreover, the suddenness of redistricting makes this a different problem than gradual migration patterns, which all states and districts will have experienced to some extent since 2020.

If 538 did not adequately control for this development – and, as far as I can tell, they [do not mention it](https://abcnews.go.com/538/538s-2024-presidential-election-forecast-works/story?id=113068753) – their embedded assumptions about NE-2 voters’ behavior could be outdated. For instance, rural voters, given their proximity to the agricultural industry, might be responsive to different economic indicators than Omahans. If that is the case, it could be that new NE-2 voters diluted the relationships that 538’s model was built on, causing the model to overpredict support from Harris. In NE-1, the same story plays out, just in reverse: NE-1 [added](https://www.politico.com/news/magazine/2024/10/14/nebraska-2nd-district-battleground-00183400) a few purple areas which previously belonged to NE-2 and dropped portions of rural northeastern Nebraska which swung dramatically to the right in 2024 (see county-by-county swing map, above), which could account for 538 underestimating support for Harris there.

## Vavreck's Campaign Typology

The redistricting explanation does not, however, rule out the possibility that the Harris campaign underperformed in Nebraska. In this section, I will argue that, although Harris did run a loosely-defined insurgent campaign in Nebraska, as recommended by [Vavreck (2009)](https://muse.jhu.edu/book/36264) for a candidate in her position, she may have chosen a non-ideal issue.

In both [public appearances](https://nebraskaexaminer.com/briefs/tim-walz-coming-back-to-nebraskas-2nd-district-for-harris-on-oct-19/) and [private meetings](https://nebraskaexaminer.com/2024/08/10/nebraskas-2nd-district-steps-back-into-presidential-spotlight-after-crazy-month/), Harris’s Nebraska surrogates emphasized reproductive rights, making abortion the frontrunner insurgent issue of her campaign in the “blue dot.” This satisfies Vavreck’s requirement that successful insurgents talk mostly about some issue other than the economy, but fares worse on her corollary conditions.

Namely, while abortion is regarded as a strong issue for Democrats nationwide, pro-choice ballot initiatives failed in Nebraska in 2024 and Representative Don Bacon, a pro-life Republican, won a Congressional seat in all-important NE-2. In other words, the public opinion on abortion may not be as lopsided in Nebraska as a strong insurgent issue should be, per Vavreck. Vavreck’s typology also states that the clarifying candidate should be hemmed in by an unpopular stance that the insurgent candidate can exploit. While Harris could point to Trump’s nomination of three conservative Supreme Court justices who joined in the decision which overturned Roe, Trump is famously hard to pin down on abortion, so it was no easy task for the Harris campaign to clearly communicate the stakes of a second Trump administration for reproductive rights.

What, then, should Harris have used as her insurgent issue in NE-2? There are at least two plausible contenders, both of which rely on a third, unstated criterion that nevertheless is essential to Vavreck's thesis: salience. Besides its weaknesses in the corollaries that Vavreck does define, abortion might fail as an insurgent issue simply because it is not top of mind for Nebraskans, several years out from the Dobbs decision. On the other hand, Social Security and candidate quality – or, as Vavreck might say, “traits” – might have helped Harris to catch up with 538’s forecast.

The Washington Examiner [conducted a Google trends analysis](https://www.washingtonexaminer.com/news/campaigns/presidential/3195453/2024-election-nebraska-voters-issues-abortion-social-security/) and reported that Social Security was a top issue for Nebraskans. I should caution that this is a simplistic method of assessing Nebraskans’ priorities that may be overinclusive of non-voters and the report is not transparent about the coding practices used (it may just be easier to code searches for a specific program, like Social Security, compared to a more generic topic like foreign policy or the economy). If, however, we assume that Social Security is a salient issue to Nebraskans, and that those who do care about the program have a lopsided preference for retaining their benefits, then the issue could have been a strong candidate for Harris’s campaign messaging. Admittedly, Trump [is once again tricky to categorize](https://www.cnn.com/2024/11/13/politics/benefits-trump-social-security-plan/index.html) on Social Security, but given Nebraskans’ pre-existing frustration with taxes on benefits in their own state under its Republican government, the Harris campaign might have been able to create some useable anxiety among Nebraska voters about what sending Trump back to the White House could do to their benefits.

Bacon – the NE-2 representative who endorsed Trump in 2024 but rejected his efforts to overrule the vote in 2020 – [suggested](https://www.politico.com/news/magazine/2024/10/14/nebraska-2nd-district-battleground-00183400) that Trump was ahead with his nonpartisan constituents on policy but behind on “trust.” National Democrats, eager to win Bacon’s seat for themselves, were unsurprisingly uninterested in building a coalition with even the moderates of Nebraska’s GOP (except, of course, when they needed Republican state legislators to shoot down winner-take-all). Perhaps this reticence contributed to shaving down Harris’s margin in NE-2, where campaign messaging focused less on abortion and more on commonsense, democratic governance – an area where Trump is almost certainly committed to unpopular positions – could have lured Trump-weary Bacon voters.

# Thank You

Thank you for joining me this semester and on this detour into Nebraskan politics!

# References

270toWin. (n.d.) Nebraska. Accessed December 9, 2024. https://www.270towin.com/states/Nebraska

Ballotpedia. (n.d.) Nebraska elections, 2024. Accessed December 7, 2024. https://ballotpedia.org/Nebraska_elections,_2024

Ballotpedia. (n.d.) Redistricting in Nebraska after the 2020 census. Accessed December 9, 2024. https://ballotpedia.org/Redistricting_in_Nebraska_after_the_2020_census

Boyd County, Nebraska. (2024, November 6). Summary Results Report. https://countyelectionresults.nebraska.gov/election_files/Boyd/2024_General_official_results.pdf

CNN Politics. (n.d.) Nebraska President Results. Accessed December 9, 2024. https://www.cnn.com/election/2020/results/state/nebraska/president

Cook Political Report. (2024, November 4). 2024 CPR Electoral College Ratings. https://www.cookpolitical.com/ratings/presidential-race-ratings

Deuel County, Nebraska. (2024, November 6). Summary Results Report. https://countyelectionresults.nebraska.gov/election_files/Deuel/2024_General_official_results.pdf

Federal Election Commission. (n.d.) Campaign finance data. Accessed December 9, 2024. https://www.fec.gov/data/

FiveThirtyEight. (2024, November 5). Who Is Favored To Win Nebraska's 1st District's 1 Electoral Vote? ABC News. https://projects.fivethirtyeight.com/2024-election-forecast/nebraska-1/

FiveThirtyEight. (2024, November 5). Who Is Favored To Win Nebraska's 2 Electoral Votes? ABC News. https://projects.fivethirtyeight.com/2024-election-forecast/nebraska/

FiveThirtyEight. (2024, November 5). Who Is Favored To Win Nebraska's 2nd District's 1 Electoral Vote? ABC News. https://projects.fivethirtyeight.com/2024-election-forecast/nebraska-2/

FiveThirtyEight. (2024, November 5). Who Is Favored To Win Nebraska's 3rd District's 1 Electoral Vote? ABC News. https://projects.fivethirtyeight.com/2024-election-forecast/nebraska-3/

Gilsinan, K. (2024, October 14). The Battleground Where Harris Is Drubbing Trump. Politico. https://www.politico.com/news/magazine/2024/10/14/nebraska-2nd-district-battleground-00183400

Hoff, M. (2024, September 12). All eyes are on Nebraska's 'blue dot,' which could determine who wins the White House. USA Today. https://www.usatoday.com/story/news/politics/elections/2024/09/12/2024-presidential-race-nebraska-blue-dot/75018361007/

Hopkins, D. (2017). Red Fighting Blue. Cambridge University Press. https://doi.org/10.1017/9781108123594.006

Kimball County, Nebraska. (2024, November 6). Kimball County General Election Results. https://irp.cdn-website.com/82f946ac/files/uploaded/General_Election_Results-Pres_2024-c4b5ed95.pdf

Koons, C. (2024, December 5). Latest 2024 farm income forecast shows overall decrease from 2023. Nebraska Examiner. https://nebraskaexaminer.com/2024/12/05/latest-2024-farm-income-forecast-shows-overall-decrease-from-2023/

Lee, C. (2024, November 6). How the 10 States' Abortion Ballot Initiatives Fared in the 2024 Election. TIME Magazine. https://time.com/7173410/abortion-ballot-results-2024-election/

Luhby, T. (2024, November 13). How a second Trump term could affect Social Security benefits. CNN Politics. https://www.cnn.com/2024/11/13/politics/benefits-trump-social-security-plan/index.html

McLoon, A. (2024, November 4). Election Week: The presidential ground game in Omaha. KETV NewsWatch 7. https://www.ketv.com/article/election-week-the-presidential-ground-game-in-omaha/62798017

Mondeaux, C. (2024, November 4). Why a Dan Osborn victory in Nebraska Senate race may not be a win for any party. Washington Examiner. https://www.washingtonexaminer.com/news/campaigns/congressional/3214917/dan-osborn-nebraska-senate-race-independent/

Morris, G. (2024, August 23). How 538's 2024 presidential election forecast works. ABC News. https://abcnews.go.com/538/538s-2024-presidential-election-forecast-works/story?id=113068753

Murray, I. (2024, October 20). 'Blue Dot' in Nebraska draws boldface political names. ABC News. https://abcnews.go.com/Politics/blue-dot-nebraska-draws-boldface-political-names/story?id=114967525

Napier, A. (2024, October 24). Election 2024: Here are the issues Nebraska voters care the most about. Washington Examiner. https://www.washingtonexaminer.com/news/campaigns/presidential/3195453/2024-election-nebraska-voters-issues-abortion-social-security/

Nebraska Secretary of State. (2024, November 26). Election Results. https://electionresults.nebraska.gov/resultsCTY.aspx?type=PC&rid=12465&osn=93&map=CTY

Nir, D. (2022, November 14). Daily Kos Elections' 2020 presidential results by congressional district. Daily Kos. https://www.dailykos.com/stories/2012/11/19/1163009/-Daily-Kos-Elections-presidential-results-by-congressional-district-for-the-2012-2008-elections

NPR. (2024, December 5). Nebraska Election Results. https://apps.npr.org/2024-election-results/nebraska.html?section=P

Sanderford, A. (2024, August 10). Nebraska's 2nd District steps back into presidential spotlight after crazy month. Nebraska Examiner. https://nebraskaexaminer.com/2024/08/10/nebraskas-2nd-district-steps-back-into-presidential-spotlight-after-crazy-month/

Sanderford, A. (2024, October 5). Gwen Walz stops in Nebraska to stump for reproductive rights, Dems. Nebraska Examiner. https://nebraskaexaminer.com/2024/10/05/gwen-walz-stops-in-nebraska-to-stump-for-reproductive-rights-dems/

Sanderford, A. (2024, October 13). Tim Walz coming back to Nebraska's 2nd District for Harris on Oct. 19. Nebraska Examiner. https://nebraskaexaminer.com/briefs/tim-walz-coming-back-to-nebraskas-2nd-district-for-harris-on-oct-19/

Sanderford, A. (2024, September 23). State Sen. Mike McDonnell deflates GOP hopes for Nebraska winner-take-all in 2024. Nebraska Examiner. https://nebraskaexaminer.com/2024/09/23/state-sen-mike-mcdonnell-deflates-gop-hopes-for-nebraska-winner-take-all-in-2024/

Schleifer, T. (2024, September 17). Presidential Campaigns and Allies Plan $500 Million in TV and Radio Ads. The New York Times. https://www.nytimes.com/2024/09/17/us/elections/presidential-campaign-advertising-spending.html

United States Census Bureau. (n.d.) Quick Facts: Nebraska. Accessed December 9, 2024. https://www.census.gov/quickfacts/fact/table/NE/PST045223

Vavreck, L. (2009). The Message Matters: The Economy and Presidential Campaigns. Princeton: Princeton University Press. https://muse.jhu.edu/book/36264
