---
title: "Campaign Narrative"
author: Ella Trembanis
date: '2024-12-07'
output:
  html_document:
    df_print: paged
categories: []
tags: []
slug: "campaign-narrative"
---

```{r load libraries, echo=FALSE, include=FALSE, warning=FALSE}
library(censable)
library(geofacet)
library(ggpubr)
library(ggthemes)
library(haven)
library(kableExtra)
library(maps)
library(mgcv)
library(mgcViz)
library(RColorBrewer)
library(readstata13)
library(scales)
library(sf)
library(spData)
library(stargazer)
library(tidygeocoder)
library(tidyverse)
library(tigris)
library(tmap)
library(tmaptools)
library(viridis)
```

```{r custom theme, echo=FALSE, include=FALSE, warning=FALSE}
my_theme <- function() {
  theme(
  # set panel features
  panel.border = element_rect(color = "#000033", fill = NA, linetype = 1),
  panel.background = element_rect(fill = "#FFFFFF"),
  panel.grid.major.x = element_line(color = "#9999CC", linetype = 2, size = 0.2),
  panel.grid.minor.x = element_blank(),
  panel.grid.major.y = element_line(color = "#9999CC", linetype = 2, size = 0.2),
  panel.grid.minor.y = element_blank(),
  # set text and axis features
  text = element_text(color = "#000033", family = "Courier"),
  title = element_text(color = "#000033", face = "bold", family = "Courier"),
  axis.ticks = element_line(color = "#000033"),
  # set legend features
  legend.position = "bottom",
  legend.background = element_rect(color = "#000033", fill = NA, linetype = 1)
  )
}
```

```{r simplified custom theme, echo=FALSE, include=FALSE, warning=FALSE}
my_map_theme <- function() {
  theme(
    # set panel features
    panel.background = element_rect(fill = "#FFFFFF"),
    # set text features
    text = element_text(color = "#000033", family = "Courier"),
    title = element_text(color = "#000033", face = "bold", family = "Courier"),
    # set legend features
    legend.position = "bottom",
    legend.title.position = "top",
    legend.background = element_rect(color = "#000033", fill = NA, linetype = 1),
    # set axis features to not display
    axis.title = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank()
  )
}
```

```{r read data, echo=FALSE, include=FALSE, warning=FALSE}
# Read 2024 results datasets. 
d_state_2024 <- read_csv("/Users/ellatrembanis/Desktop/Gov 1347/election-blog/election-blog/Week10/data/state_votes_pres_2024.csv")[-1, 1:6]
d_county_2024 <- read_csv("/Users/ellatrembanis/Desktop/Gov 1347/election-blog/election-blog/Week10/data/county_votes_pres_2024.csv")[-1, 1:6]
d_county_2020 <- read_csv("/Users/ellatrembanis/Desktop/Gov 1347/election-blog/election-blog/Week10/data/county_votes_pres_2020.csv")[-1, 1:6]

# Process 2024 state and county-level data. 
d_state_2024 <- d_state_2024 |> 
  mutate(FIPS = as.numeric(FIPS), 
         votes_trump = as.numeric(`Donald J. Trump`), 
         votes_harris = as.numeric(`Kamala D. Harris`), 
         votes = as.numeric(`Total Vote`), 
         trump_pv = votes_trump/votes, 
         harris_pv = votes_harris/votes, 
         trump_2pv = votes_trump/(votes_trump + votes_harris), 
         harris_2pv = votes_harris/(votes_trump + votes_harris)) |> 
  mutate(winner = case_when(votes_trump > votes_harris ~ "REP", 
                            .default = "DEM")) |> 
  select(FIPS, `Geographic Name`, `Geographic Subtype`, votes_trump, votes_harris, votes, 
         winner, trump_pv, harris_pv, trump_2pv, harris_2pv)

d_county_2024 <- d_county_2024 |>
  mutate(FIPS = as.numeric(FIPS),
         votes_trump = as.numeric(`Donald J. Trump`), 
         votes_harris = as.numeric(`Kamala D. Harris`), 
         votes = as.numeric(`Total Vote`), 
         trump_pv = votes_trump/votes, 
         harris_pv = votes_harris/votes, 
         trump_2pv = votes_trump/(votes_trump + votes_harris), 
         harris_2pv = votes_harris/(votes_trump + votes_harris)) |> 
  mutate(winner = case_when(votes_trump > votes_harris ~ "REP", 
                            .default = "DEM")) |> 
  select(FIPS, `Geographic Name`, `Geographic Subtype`, votes_trump, votes_harris, votes, 
         winner, trump_pv, harris_pv, trump_2pv, harris_2pv)

d_county_2020 <- d_county_2020 |> 
  mutate(FIPS = as.numeric(FIPS),
         votes_trump_2020 = as.numeric(`Donald J. Trump`), 
         votes_biden_2020 = as.numeric(`Joseph R. Biden Jr.`), 
         votes_2020 = as.numeric(`Total Vote`), 
         trump_pv_2020 = votes_trump_2020/votes_2020, 
         biden_pv_2020 = votes_biden_2020/votes_2020, 
         trump_2pv_2020 = votes_trump_2020/(votes_trump_2020 + votes_biden_2020), 
         biden_2pv_2020 = votes_biden_2020/(votes_trump_2020 + votes_biden_2020)) |> 
  mutate(winner_2020 = case_when(votes_trump_2020 > votes_biden_2020 ~ "REP", 
                            .default = "DEM")) |> 
  select(FIPS, `Geographic Name`, `Geographic Subtype`, votes_trump_2020, votes_biden_2020, votes_2020, 
         winner_2020, trump_pv_2020, biden_pv_2020, trump_2pv_2020, biden_2pv_2020)
```

```{r map set-up, echo=FALSE, include=FALSE, warning=FALSE}
# Sequester state and county-level map.
states_2024 <- states(cb = TRUE, year = 2023) |> 
  shift_geometry() |> 
  mutate(GEOID = as.numeric(GEOID)) |> 
  left_join(d_state_2024, by = c("GEOID" = "FIPS")) |> 
  drop_na()
counties_2024 <- counties(cb = TRUE, resolution = "5m", year = 2023) |> 
  shift_geometry() |> 
  mutate(GEOID = as.numeric(GEOID)) |> 
  left_join(d_county_2024, by = c("GEOID" = "FIPS")) |> 
  left_join(d_county_2020, by = c("GEOID" = "FIPS")) |>
  mutate(shift = (trump_pv - trump_pv_2020) * 100, 
         shift_dir = case_when(shift > 0 ~ "REP", 
                               shift < 0 ~ "DEM", 
                               TRUE ~ "No Change"),
         centroid = st_centroid(geometry), 
         centroid_long = st_coordinates(centroid)[,1],
         centroid_lat = st_coordinates(centroid)[,2],
         scale_factor = 1e4, 
         end_long = centroid_long + scale_factor * shift,
         end_lat = centroid_lat + scale_factor * shift) |>
  drop_na()
county_pop_2024 <- read_csv("/Users/ellatrembanis/Desktop/Gov 1347/election-blog/election-blog/Week10/data/PopulationEstimates.csv") |> 
  mutate(FIPStxt = as.numeric(FIPStxt)) |>
  select(FIPStxt, POP_ESTIMATE_2023)
counties_2024 <- counties_2024 |> 
  left_join(county_pop_2024, by = c("GEOID" = "FIPStxt"))
```

```{r ne table code, echo=FALSE, include=FALSE, warning=FALSE}
ne_results <- tibble(`Unit of Analysis` = c("Statewide", "CD-1", "CD-2", "CD-3"),
                     "Harris Two-Party Vote Share (2024)" = c(39.4, 43.4, 52.3, 22.8),
                     "My Prediction" = c("40.2", "NA", "NA", "NA"),
                     "538's Prediction" = c(41.3, 41.6, 53.5, 22.6),
                     "CPR Rating" = c("Solid R", "Solid R", "Likely D", "Solid R"),
                     "Biden Two-Party Vote Share (2020)" = c(39.2,43.3, 52.2, 23.1))

ne_table <- knitr::kable(ne_results)
```

```{r ne map code, echo=FALSE, include=FALSE, warning=FALSE}
# Check county-level shifts in Nebraska between 2024 and 2020. 
ne_map <- counties_2024 |> 
  filter(STATE_NAME == "Nebraska") |> 
  ggplot() +
  geom_sf(fill = "#FCFCFC", color = "#000033") +  # Base map
  geom_label(aes(x = centroid_long, y = centroid_lat-1e4, label = NAME),
            size = 2,  # Adjust size as needed
            color = "black", hjust = 0.5, vjust = -0.005) + 
  geom_curve(aes(x = centroid_long, 
                 y = centroid_lat,
                 xend = end_long, 
                 yend = end_lat,
                 color = shift_dir),
             arrow = arrow(length = unit(0.1, "cm"), type = "closed"),  # Smaller arrowhead
             curvature = 0.2,  # Add a slight curve to each arrow
             size = 0.3) +
  scale_color_manual(values = c("DEM" = "#7CB9E8", "REP" = "#FF6680")) +
  my_map_theme() +
  labs(title = "Presidential Voting Shifts by County (2020-2024)")
```

# Campaigning in the Cornhusker State



```{r print ne table, echo=FALSE, include=TRUE, warning=FALSE}
ne_table
```

```{r print ne map, echo=FALSE, include=TRUE, warning=FALSE}
ne_map
```

# Nebraska: An Overview

## Demographics, Geography, & Electoral History

## Red State, Purple Dot

No doubt the most famous feature of Nebraska’s electoral system is its method of splitting its five electoral votes. Unlike most states, which award all of their votes to whichever candidate receives a plurality in the state-level popular vote, Nebraska awards just two votes to the statewide winner and one vote to the winner of each of its three congressional districts.

Compared to winner-take-all, this configuration is a boon for Democrats; in 2008 and 2020, Nebraska’s second congressional district awarded one vote to the Democratic candidate. However, the idea that NE-2 is a reliable “blue dot” in the ruby-red state is at least a little overstated. This is a district, after all, that voted for a Republican presidential candidate as recently as [2016](https://nebraskaexaminer.com/2024/08/10/nebraskas-2nd-district-steps-back-into-presidential-spotlight-after-crazy-month/) – a district that, to this day, is represented in the House by a moderate Republican. Even in [Omaha](https://www.politico.com/news/magazine/2024/10/14/nebraska-2nd-district-battleground-00183400), the city nominally driving NE-2’s “blue dot” status, the mayor is a Republican and registered Republicans outnumber registered Democrats. As the national electorate calcifies, and “red” and “blue” states become steadily more predictable, the so-called “blue dot” has more in common with the kind of Midwestern battleground that [Hopkins](https://doi.org/10.1017/9781108123594.006) described in the wake of the 2016 election: purple, elastic, and torn between urban and rural constituents.

Earlier this fall, Nebraska Republicans, with Trump’s blessing, [launched](https://nebraskaexaminer.com/2024/08/10/nebraskas-2nd-district-steps-back-into-presidential-spotlight-after-crazy-month/) a campaign to re-institute winner-take-all. The push [fell short](https://nebraskaexaminer.com/2024/09/23/state-sen-mike-mcdonnell-deflates-gop-hopes-for-nebraska-winner-take-all-in-2024/) in the state legislature thanks to a pivotal “no” vote from Republican Mike McDonnell, a former Omaha Democrat. McDonnell, however, did not set aside the possibility of changing Nebraska’s electoral vote allocation method in time for the next presidential election.

## Elsewhere on the Ballot

Of course, the presidential campaign did not happen in isolation. Nebraska voters faced other important down-ballot choices on Election Day, and the state generally affirmed its preference for Republican officeholders and conservative policies.

Republican Senator Deb Fischer was [re-elected](https://ballotpedia.org/United_States_Senate_election_in_Nebraska,_2024) over Independent challenger Dan Osborn. Osborn’s middle-of-the-road campaign, which emphasized his nonpartisanship and his working-class roots, made a reasonably competitive race out of what might otherwise have been a slam-dunk for Fischer. There are difficulties in drawing analogies between this race and the presidential contest, since both candidates [attempted](https://www.washingtonexaminer.com/news/campaigns/congressional/3214917/dan-osborn-nebraska-senate-race-independent/) to align themselves with Trump – though only Fischer received his coveted endorsement.

Nebraskans also weighed in on a range of ballot initiatives, and the results are a fascinating mixed bag. Balking the post-Dobbs [trend](https://time.com/7173410/abortion-ballot-results-2024-election/) of enthusiasm for reproductive rights on the ballot, Nebraskans [struck down](https://ballotpedia.org/Nebraska_elections,_2024) a measure to enshrine a right to abortion in the state constitution and supported a ban for abortions after the first trimester of pregnancy. Interestingly, the margin was substantially lower for the former initiative (3 points versus 10 points) and fewer voters responded to the latter question. Perhaps it was easier to mobilize voters around the relatively clear terms of the ban – which specifies the trimester of concern as well as exceptions for rape, incest, and the health of the mother – compared to the somewhat more vague language of “fetal viability” used in the proposed constitutional amendment.

Nevertheless, voters also opted to require paid sick leave, to legalize medical marijuana (and establish a regulatory body to oversee its usage), and to overturn a law which would provide vouchers to qualified private school students.

Finally, voters in the state’s all-important second Congressional District [chose](https://ballotpedia.org/Nebraska%27s_2nd_Congressional_District_election,_2024) to return Republican incumbent Don Bacon to the House of Representatives. As a testament to the purple-ness of the district, however, Bacon’s opponent, Democrat Tony Vargas, lost by a slim 1.8-point margin.

# 2024 Campaign Activity



# Conclusions



# References

Ballotpedia. (n.d.) Nebraska elections, 2024. Accessed December 7, 2024. https://ballotpedia.org/Nebraska_elections,_2024

CNN Politics. (n.d.) Nebraska President Results. Accessed December 9, 2024. https://www.cnn.com/election/2020/results/state/nebraska/president

Cook Political Report. (2024, November 4). 2024 CPR Electoral College Ratings. https://www.cookpolitical.com/ratings/presidential-race-ratings

Deuel County, Nebraska. (2024, November 6). Summary Results Report. https://countyelectionresults.nebraska.gov/election_files/Deuel/2024_General_official_results.pdf

Federal Election Commission. (n.d.) Campaign finance data. Accessed December 9, 2024. https://www.fec.gov/data/

FiveThirtyEight. (2024, November 5). Who Is Favored To Win Nebraska's 1st District's 1 Electoral Vote? ABC News. https://projects.fivethirtyeight.com/2024-election-forecast/nebraska-1/

FiveThirtyEight. (2024, November 5). Who Is Favored To Win Nebraska's 2 Electoral Votes? ABC News. https://projects.fivethirtyeight.com/2024-election-forecast/nebraska/

FiveThirtyEight. (2024, November 5). Who Is Favored To Win Nebraska's 2nd District's 1 Electoral Vote? ABC News. https://projects.fivethirtyeight.com/2024-election-forecast/nebraska-2/

FiveThirtyEight. (2024, November 5). Who Is Favored To Win Nebraska's 3rd District's 1 Electoral Vote? ABC News. https://projects.fivethirtyeight.com/2024-election-forecast/nebraska-3/

Gilsinan, K. (2024, October 14). The Battleground Where Harris Is Drubbing Trump. Politico. https://www.politico.com/news/magazine/2024/10/14/nebraska-2nd-district-battleground-00183400

Hoff, M. (2024, September 12). All eyes are on Nebraska's 'blue dot,' which could determine who wins the White House. USA Today. https://www.usatoday.com/story/news/politics/elections/2024/09/12/2024-presidential-race-nebraska-blue-dot/75018361007/

Hopkins, D. (2017). Red Fighting Blue. Cambridge University Press. https://doi.org/10.1017/9781108123594.006

Kimball County, Nebraska. (2024, November 6). Kimball County General Election Results. https://irp.cdn-website.com/82f946ac/files/uploaded/General_Election_Results-Pres_2024-c4b5ed95.pdf

Lee, C. (2024, November 6). How the 10 States' Abortion Ballot Initiatives Fared in the 2024 Election. TIME Magazine. https://time.com/7173410/abortion-ballot-results-2024-election/

McLoon, A. (2024, November 4). Election Week: The presidential ground game in Omaha. KETV NewsWatch 7. https://www.ketv.com/article/election-week-the-presidential-ground-game-in-omaha/62798017

Mondeaux, C. (2024, November 4). Why a Dan Osborn victory in Nebraska Senate race may not be a win for any party. Washington Examiner. https://www.washingtonexaminer.com/news/campaigns/congressional/3214917/dan-osborn-nebraska-senate-race-independent/

Murray, I. (2024, October 20). 'Blue Dot' in Nebraska draws boldface political names. ABC News. https://abcnews.go.com/Politics/blue-dot-nebraska-draws-boldface-political-names/story?id=114967525

Napier, A. (2024, October 24). Election 2024: Here are the issues Nebraska voters care the most about. Washington Examiner. https://www.washingtonexaminer.com/news/campaigns/presidential/3195453/2024-election-nebraska-voters-issues-abortion-social-security/

Nir, D. (2022, November 14). Daily Kos Elections' 2020 presidential results by congressional district. Daily Kos. https://www.dailykos.com/stories/2012/11/19/1163009/-Daily-Kos-Elections-presidential-results-by-congressional-district-for-the-2012-2008-elections

NPR. (2024, December 5). Nebraska Election Results. https://apps.npr.org/2024-election-results/nebraska.html?section=P

Sanderford, A. (2024, August 10). Nebraska's 2nd District steps back into presidential spotlight after crazy month. Nebraska Examiner. https://nebraskaexaminer.com/2024/08/10/nebraskas-2nd-district-steps-back-into-presidential-spotlight-after-crazy-month/

Sanderford, A. (2024, September 23). State Sen. Mike McDonnell deflates GOP hopes for Nebraska winner-take-all in 2024. Nebraska Examiner. https://nebraskaexaminer.com/2024/09/23/state-sen-mike-mcdonnell-deflates-gop-hopes-for-nebraska-winner-take-all-in-2024/

Schleifer, T. (2024, September 17). Presidential Campaigns and Allies Plan $500 Million in TV and Radio Ads. The New York Times. https://www.nytimes.com/2024/09/17/us/elections/presidential-campaign-advertising-spending.html
