---
title: "Blog #2 Code"
author: "Ella Trembanis"
date: "2024-09-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Set-Up

```{r load packages, echo=FALSE}
library(ggplot2)
library(tidyverse)
```

```{r custom theme, echo=FALSE}
my_theme <- function() {
  theme(
  # set panel features
  panel.border = element_rect(color = "#000033", fill = NA, linetype = 1),
  panel.background = element_rect(fill = "#FFFFFF"),
  panel.grid.major.x = element_line(color = "#9999CC", linetype = 2, size = 0.2),
  panel.grid.minor.x = element_blank(),
  panel.grid.major.y = element_line(color = "#9999CC", linetype = 2, size = 0.2),
  panel.grid.minor.y = element_blank(),
  # set text and axis features
  text = element_text(color = "#000033", family = "Courier"),
  title = element_text(color = "#000033", face = "bold", family = "Courier"),
  axis.ticks = element_line(color = "#000033"),
  # set legend features
  legend.position = "bottom",
  legend.background = element_rect(color = "#000033", fill = NA, linetype = 1)
  )
}
```

# Load popular vote data
```{r load pop}
popvote <- read_csv("/Users/ellatrembanis/Desktop/Gov 1347/election-blog/election-blog/Week2/data/popvote_1948-2020.csv")
```

# FRED variables:

# GDP, billions $, 1947-2024
# GDP_growth_quarterly, %
# RDPI, $, 1959-2024
# RDPI_growth_quarterly, %
# CPI, $ index, 1947-2024
# unemployment, %, 1948-2024
# sp500_, $, 1927-2024 

```{r load fred}
fred <- read_csv("/Users/ellatrembanis/Desktop/Gov 1347/election-blog/election-blog/Week2/data/fred_econ.csv")
```

# BEA variables:

# GDP, 1947-2024 (all)
# GNP
# RDPI
# Personal consumption expenditures
# Goods
# Durable goods
# Nondurable goods
# Services 
# Population (midperiod, thousands)

```{r load bea}
bea <- read_csv("/Users/ellatrembanis/Desktop/Gov 1347/election-blog/election-blog/Week2/data/bea_econ.csv") |> 
  # standardize column names
  rename(year = "Year",
         quarter = "Quarter", 
         gdp = "Gross domestic product", 
         gnp = "Gross national product", 
         dpi = "Disposable personal income", 
         consumption = "Personal consumption expenditures", 
         goods = "Goods", 
         durables = "Durable goods", 
         nondurables = "Nondurable goods", 
         services = "Services", 
         pop = "Population (midperiod, thousands)")
```

# Filter and Merge

```{r clean up}
party_econ <- popvote |> 
  # incumbent party vote share only
  filter(incumbent_party == TRUE) |> 
  select(year, pv, pv2p, winner, incumbent) |> 
  # filter to Q2 (most recent available quarter of current year)
  left_join(fred |> filter(quarter == 2)) |> 
  left_join(bea |> filter(quarter == "Q2") |> 
              select(year, dpi))

inc_econ <- popvote |>
  # incumbent (individual) vote share only, omit 2020
  filter(incumbent == TRUE & year != 2020) |>
  select(year, pv, pv2p, winner) |>
  left_join(fred |> 
              filter(quarter == 2)) |>
  left_join(bea |> 
              filter(quarter == "Q2") |>
              select(year, dpi))

heir_econ <- popvote |>
  # incumbent (same-party heir) vote share only
  filter(incumbent_party == TRUE & incumbent == FALSE) |>
  select(year, pv, pv2p, winner) |>
  left_join(fred |>
              filter(quarter == 2)) |>
  left_join(bea |>
              filter(quarter == "Q2") |>
              select(year, dpi))
```

# Assess GDP

```{r gdp scatter w/ 2020}
party_econ |> 
  ggplot(aes(x = GDP_growth_quarterly, y = pv2p, label = year)) + 
  geom_label(aes(family = "Courier")) + 
  geom_smooth(method = "lm", formula = y ~ x, color = "#9999CC") +
  geom_hline(yintercept = 50, lty = 2, color = "#000033") + 
  geom_vline(xintercept = 0.01, lty = 2, color = "#000033") + 
  labs(x = "Second Quarter GDP Growth (%)", 
       y = "Popular Vote Share - Incumbent Party (%)", 
       title = "Fig. I - Q2 GDP vs. Vote Share") + 
  my_theme() + 
  theme(plot.title = element_text(size = 18))
```

```{r gdp scatter w/o outliers}
party_econ_omit <- party_econ |>
  filter(year != 2020 & year != 1980)

party_econ_omit |> 
  ggplot(aes(x = GDP_growth_quarterly, y = pv2p, label = year)) + 
  geom_label(aes(family = "Courier")) + 
  geom_smooth(method = "lm", formula = y ~ x, color = "#9999CC") +
  geom_hline(yintercept = 50, lty = 2, color = "#000033") + 
  geom_vline(xintercept = 0.01, lty = 2, color = "#000033") + 
  labs(x = "Second Quarter GDP Growth (%)", 
       y = "Popular Vote Share - Incumbent Party (%)", 
       title = "Fig. II - Q2 GDP vs. Vote Share, No Outliers") + 
  my_theme() + 
  theme(plot.title = element_text(size = 18))
```

```{r gdp boxplot}
ggplot(party_econ, aes(x=GDP_growth_quarterly)) +
  geom_boxplot() +
  xlab("Second Quarter GDP Growth (%)") +
  ggtitle("Q2 GDP Growth Boxplot") +
  geom_label(x=-27.9, y=0.05, label = "2020") +
  geom_label(x=-8, y=0.05, label = "1980") +
  my_theme() +
  theme(plot.title = element_text(size = 18))
```

# Fit bivariate OLS
```{r OLS gdp}
gdp_with <- lm(pv2p ~ GDP_growth_quarterly, 
               data = party_econ)
gdp_with |> summary()

gdp_without <- lm(pv2p ~ GDP_growth_quarterly, 
                         data = party_econ_omit)
gdp_without |> summary()
```

# Compute correlations between Q2 GDP growth and incumbent vote 2-party vote share.
```{r correlation}
cor(party_econ$GDP_growth_quarterly, 
    party_econ$pv2p) # with 2020

cor(party_econ_omit$GDP_growth_quarterly, 
    party_econ_omit$pv2p) # drop 2020
```

# Assess unemployment

```{r unemployment scatter}
party_econ |> 
  ggplot(aes(x = unemployment, y = pv2p, label = year)) + 
  geom_label(aes(family = "Courier")) + 
  geom_smooth(method = "lm", formula = y ~ x, color = "#9999CC") +
  geom_hline(yintercept = 50, lty = 2, color = "#000033") + 
  geom_vline(xintercept = 0.01, lty = 2, color = "#000033") + 
  labs(x = "Second Quarter Unemployment Rate (%)", 
       y = "Popular Vote Share - Incumbent Party (%)", 
       title = "Fig. X - Q2 Unemployment vs. Vote Share") + 
  my_theme() + 
  theme(plot.title = element_text(size = 18))
```

```{r unemployment scatter w/o 2020}
party_econ_2020 <- party_econ |>
  filter(year != 2020)

party_econ_2020 |> 
  ggplot(aes(x = unemployment, y = pv2p, label = year)) + 
  geom_label(aes(family = "Courier")) + 
  geom_smooth(method = "lm", formula = y ~ x, color = "#9999CC") +
  geom_hline(yintercept = 50, lty = 2, color = "#000033") + 
  geom_vline(xintercept = 0.01, lty = 2, color = "#000033") + 
  labs(x = "Second Quarter Unemployment Rate (%)", 
       y = "Popular Vote Share - Incumbent Party (%)", 
       title = "Fig. III - Q2 Unemployment vs. Vote Share, 2020 Omitted") + 
  my_theme() + 
  theme(plot.title = element_text(size = 18))
```

```{r OLS unemployment}
unemp_with <- lm(pv2p ~ unemployment, 
               data = party_econ)
unemp_with |> summary()

unemp_without <- lm(pv2p ~ unemployment, 
                         data = party_econ_2020)
unemp_without |> summary()
```

```{r unemp boxplot}
ggplot(party_econ, aes(x=unemployment)) +
  geom_boxplot() +
  xlab("Second Quarter Unemployment Rate (%)") +
  geom_label(x=13, y=0.05, label = "2020") +
  ggtitle("Q2 Unemployment Boxplot") +
  my_theme() +
  theme(plot.title = element_text(size = 18))
```

```{r}
gdp_without_res <- resid(gdp_without)
plot(fitted(gdp_without), gdp_without_res)
abline(0,0)
```

# Assess DPI

```{r dpi scatter}
party_econ |> 
  ggplot(aes(x = dpi, y = pv2p, label = year)) + 
  geom_label(aes(family = "Courier")) + 
  geom_smooth(method = "lm", formula = y ~ x, color = "#9999CC") +
  geom_hline(yintercept = 50, lty = 2, color = "#000033") + 
  geom_vline(xintercept = 0.01, lty = 2, color = "#000033") + 
  labs(x = "Second Quarter Disposable Personal Income", 
       y = "Popular Vote Share - Incumbent Party (%)", 
       title = "Fig. X - Q2 DPI vs. Vote Share") + 
  my_theme() + 
  theme(plot.title = element_text(size = 18))
```

```{r OLS dpi}
dpi_with <- lm(pv2p ~ dpi, 
               data = party_econ)
dpi_with |> summary()

dpi_without <- lm(pv2p ~ dpi, 
                         data = party_econ_omit)
dpi_without |> summary()
```

```{r dpi boxplot}
ggplot(party_econ, aes(x=dpi)) +
  geom_boxplot() +
  xlab("Second Quarter Disposable Personal Income") +
  ggtitle("Q2 DPI Boxplot") +
  my_theme() +
  theme(plot.title = element_text(size = 18))
```

# Assess CPI

```{r cpi scatter}
party_econ |> 
  ggplot(aes(x = CPI, y = pv2p, label = year)) + 
  geom_label(aes(family = "Courier")) + 
  geom_smooth(method = "lm", formula = y ~ x, color = "#9999CC") +
  geom_hline(yintercept = 50, lty = 2, color = "#000033") + 
  geom_vline(xintercept = 0.01, lty = 2, color = "#000033") + 
  labs(x = "Second Quarter Consumer Price Index", 
       y = "Popular Vote Share - Incumbent Party (%)", 
       title = "Fig. IV - Q2 CPI vs. Vote Share") + 
  my_theme() + 
  theme(plot.title = element_text(size = 18))
```

```{r OLS cpi}
cpi_with <- lm(pv2p ~ CPI, 
               data = party_econ)
cpi_with |> summary()

cpi_without <- lm(pv2p ~ CPI, 
                         data = party_econ_omit)
cpi_without |> summary()
```

```{r cpi boxplot}
ggplot(party_econ, aes(x=CPI)) +
  geom_boxplot() +
  xlab("Second Quarter Consumer Price Index") +
  ggtitle("Q2 CPI Boxplot") +
  my_theme() +
  theme(plot.title = element_text(size = 18))
```

# Multiple OLS

```{r multi OLS}
multi_with <- lm(pv2p ~ GDP_growth_quarterly + unemployment + CPI, 
               data = party_econ)
multi_with |> summary()

multi_without <- lm(pv2p ~ GDP_growth_quarterly + unemployment + CPI,
                    data = party_econ_omit)
multi_without |> summary()
```

# A Tale of Two Incumbents

```{r}
party_econ_omit |> 
  ggplot(aes(x = GDP_growth_quarterly, y = pv2p, label = year, fill = incumbent)) + 
  geom_label(aes(family = "Courier")) + 
  geom_hline(yintercept = 50, lty = 2, color = "#000033") + 
  geom_vline(xintercept = 0.01, lty = 2, color = "#000033") + 
  labs(x = "Second Quarter GDP Growth (%)", 
       y = "Popular Vote Share - Incumbent Party (%)", 
       title = "Fig. V - Q2 GDP vs. Vote Share by Incumbency") + 
  my_theme() + 
  theme(plot.title = element_text(size = 18))
```

```{r incumbency ols}
gdp_inc <- lm(pv2p ~ GDP_growth_quarterly,
              data = inc_econ)
gdp_inc |> summary()

gdp_heir <- lm(pv2p ~ GDP_growth_quarterly, 
               data = heir_econ)
gdp_heir |> summary()
```

```{r incumbent unemp scatter}
party_econ_omit |> 
  ggplot(aes(x = unemployment, y = pv2p, label = year, fill = incumbent)) + 
  geom_label(aes(family = "Courier")) + 
  geom_smooth(method = "lm", formula = y ~ x, color = "#9999CC") +
  geom_hline(yintercept = 50, lty = 2, color = "#000033") + 
  geom_vline(xintercept = 0.01, lty = 2, color = "#000033") + 
  labs(x = "Second Quarter Unemployment Rate (%)", 
       y = "Popular Vote Share - Incumbent Party (%)", 
       title = "Fig. VI - Q2 Unemployment vs. Vote Share, 2020 Omitted") + 
  my_theme() + 
  theme(plot.title = element_text(size = 18))
```

```{r}
unemp_inc <- lm(pv2p ~ unemployment,
              data = inc_econ)
unemp_inc |> summary()

unemp_heir <- lm(pv2p ~ unemployment, 
               data = heir_econ)
unemp_heir |> summary()
```


# 2024 Prediction

# Sequester 2024 data
```{r sequester 2024}
GDP_new <- fred |> 
  filter(year == 2024 & quarter == 2) |> 
  select(GDP_growth_quarterly)

unemp_new <- fred |>
  filter(year == 2024 & quarter == 2) |>
  select(unemployment)

dpi_new <- bea |>
  filter(year == 2024 & quarter == "Q2") |>
  select(dpi)

cpi_new <- fred |>
  filter(year == 2024 & quarter == 2) |>
  select(CPI)

multi_new <- c(GDP_new, unemp_new, dpi_new, cpi_new)
```

```{r cross-validation 1k runs}
out_samp_errors <- sapply(1:1000, function(i) {
  years_out_samp <- sample(party_econ_omit$year, 9)
  mod <- lm(pv2p ~ GDP_growth_quarterly,
            party_econ_omit[!(party_econ_omit$year %in% years_out_samp),])
  out_samp_pred <- predict(mod, party_econ_omit[party_econ_omit$year %in% years_out_samp,])
  out_samp_truth <- party_econ_omit$pv2p[party_econ_omit$year %in% years_out_samp]
  mean(out_samp_pred - out_samp_truth)
})
```

```{r}
hist(x = out_samp_errors)
```

```{r mean out-of-sample residuals}
mean(abs(out_samp_errors))
```

```{r compare to pv2p range}
range(party_econ_omit$pv2p)
```

# Predict
```{r predict w/ diff models}
# use q2 gdp for 2024 to predict next pop vote (2 outliers omitted)
predict(gdp_without, GDP_new, interval = "prediction")

# use q2 unemployment for 2024 to predict next pop vote (1 outlier omitted)
predict(unemp_without, unemp_new, interval = "prediction")

# use q2 cpi for 2024 to predict next pop vote
predict(cpi_with, cpi_new, interval = "prediction")

# use multi OLS to predict next pop vote (2 outliers omitted)
predict(multi_without, multi_new, interval = "prediction")
```

```{r predictions table}
predictions <- tibble(Model = c("GDP Growth", "Unemployment", "CPI", "Multiple"),
                 Prediction = c(51.35652, 51.97273, 48.33048, 46.30151),
                 `Lower Bound` = c(41.25755, 39.83306, 35.591, 33.24755),
                 `Upper Bound` = c(61.4555, 64.11241, 61.06995, 59.35547))
knitr::kable(predictions, digits = 2)
# Note: multiple regression prediction excludes 1980 & 2020, since these are outliers in terms of GDP growth, the only significant variable. The multiple regression which includes these points had similar downsides (see multi ols code chunk for a detailed regression output). That different predictors identify different outliers is another difficulty of the multivariate approach.
```

```{r}
predict(gdp_without, GDP_new, interval = "prediction", level = 0.8)
```

