---
title: "Blog #2 Code"
author: "Ella Trembanis"
date: "2024-09-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Set-Up

```{r load packages, echo=FALSE}
library(ggplot2)
library(tidyverse)
```

```{r custom theme, echo=FALSE}
my_theme <- function() {
  theme(
  # set panel features
  panel.border = element_rect(color = "#000033", fill = NA, linetype = 1),
  panel.background = element_rect(fill = "#FFFFFF"),
  panel.grid.major.x = element_line(color = "#9999CC", linetype = 2, size = 0.2),
  panel.grid.minor.x = element_blank(),
  panel.grid.major.y = element_line(color = "#9999CC", linetype = 2, size = 0.2),
  panel.grid.minor.y = element_blank(),
  # set text and axis features
  text = element_text(color = "#000033", family = "Courier"),
  title = element_text(color = "#000033", face = "bold", family = "Courier"),
  axis.ticks = element_line(color = "#000033"),
  # set legend features
  legend.position = "bottom",
  legend.background = element_rect(color = "#000033", fill = NA, linetype = 1)
  )
}
```

# Load popular vote data
```{r load pop}
popvote <- read_csv("/Users/ellatrembanis/Desktop/Gov 1347/election-blog/election-blog/Week2/data/popvote_1948-2020.csv")
```

# FRED variables:

# GDP, billions $, 1947-2024
# GDP_growth_quarterly, %
# RDPI, $, 1959-2024
# RDPI_growth_quarterly, %
# CPI, $ index, 1947-2024
# unemployment, %, 1948-2024
# sp500_, $, 1927-2024 

```{r load fred}
fred <- read_csv("/Users/ellatrembanis/Desktop/Gov 1347/election-blog/election-blog/Week2/data/fred_econ.csv")
```

# BEA variables:

# GDP, 1947-2024 (all)
# GNP
# RDPI
# Personal consumption expenditures
# Goods
# Durable goods
# Nondurable goods
# Services 
# Population (midperiod, thousands)

```{r load bea}
bea <- read_csv("/Users/ellatrembanis/Desktop/Gov 1347/election-blog/election-blog/Week2/data/bea_econ.csv") |> 
  # standardize column names
  rename(year = "Year",
         quarter = "Quarter", 
         gdp = "Gross domestic product", 
         gnp = "Gross national product", 
         dpi = "Disposable personal income", 
         consumption = "Personal consumption expenditures", 
         goods = "Goods", 
         durables = "Durable goods", 
         nondurables = "Nondurable goods", 
         services = "Services", 
         pop = "Population (midperiod, thousands)")
```

# Filter and Merge

```{r clean up}
inc_econ <- popvote |> 
  # incumbent vote share only
  filter(incumbent_party == TRUE) |> 
  select(year, pv, pv2p, winner) |> 
  # filter to Q2 (most recent available quarter of current year)
  left_join(fred |> filter(quarter == 2)) |> 
  left_join(bea |> filter(quarter == "Q2") |> 
              select(year, dpi))
```

# Assess GDP

```{r gdp scatter w/ 2020}
inc_econ |> 
  ggplot(aes(x = GDP_growth_quarterly, y = pv2p, label = year)) + 
  geom_label(aes(family = "Courier")) + 
  geom_smooth(method = "lm", formula = y ~ x, color = "#9999CC") +
  geom_hline(yintercept = 50, lty = 2, color = "#000033") + 
  geom_vline(xintercept = 0.01, lty = 2, color = "#000033") + 
  labs(x = "Second Quarter GDP Growth (%)", 
       y = "Popular Vote Share - Incumbent Party (%)", 
       title = "Fig. I - Q2 GDP vs. Vote Share") + 
  my_theme() + 
  theme(plot.title = element_text(size = 18))
```

```{r gdp scatter w/o 2020}
inc_econ_omit <- inc_econ |>
  filter(year != 2020)

inc_econ_omit |> 
  ggplot(aes(x = GDP_growth_quarterly, y = pv2p, label = year)) + 
  geom_label(aes(family = "Courier")) + 
  geom_smooth(method = "lm", formula = y ~ x, color = "#9999CC") +
  geom_hline(yintercept = 50, lty = 2, color = "#000033") + 
  geom_vline(xintercept = 0.01, lty = 2, color = "#000033") + 
  labs(x = "Second Quarter GDP Growth (%)", 
       y = "Popular Vote Share - Incumbent Party (%)", 
       title = "Fig. II - Q2 GDP vs. Vote Share, 2020 Omitted") + 
  my_theme() + 
  theme(plot.title = element_text(size = 18))
```

```{r gdp boxplot}
ggplot(inc_econ, aes(x=GDP_growth_quarterly)) +
  geom_boxplot() +
  xlab("Second Quarter GDP Growth (%)") +
  ggtitle("Q2 GDP Growth Boxplot") +
  geom_label(x=-27.9, y=0.05, label = "2020") +
  geom_label(x=-8, y=0.05, label = "1980") +
  my_theme() +
  theme(plot.title = element_text(size = 18))
```

# Fit bivariate OLS
```{r OLS gdp}
gdp_with <- lm(pv2p ~ GDP_growth_quarterly, 
               data = inc_econ)
gdp_with |> summary()

gdp_without <- lm(pv2p ~ GDP_growth_quarterly, 
                         data = inc_econ_omit)
gdp_without |> summary()
```

# Compute correlations between Q2 GDP growth and incumbent vote 2-party vote share.
```{r correlation}
cor(inc_econ$GDP_growth_quarterly, 
    inc_econ$pv2p) # with 2020

cor(inc_econ_omit$GDP_growth_quarterly, 
    inc_econ_omit$pv2p) # drop 2020
```

# Assess unemployment

```{r unemployment scatter}
inc_econ |> 
  ggplot(aes(x = unemployment, y = pv2p, label = year)) + 
  geom_label(aes(family = "Courier")) + 
  geom_smooth(method = "lm", formula = y ~ x, color = "#9999CC") +
  geom_hline(yintercept = 50, lty = 2, color = "#000033") + 
  geom_vline(xintercept = 0.01, lty = 2, color = "#000033") + 
  labs(x = "Second Quarter Unemployment Rate (%)", 
       y = "Popular Vote Share - Incumbent Party (%)", 
       title = "Fig. III - Q2 Unemployment vs. Vote Share") + 
  my_theme() + 
  theme(plot.title = element_text(size = 18))
```

```{r unemployment scatter w/o 2020}
inc_econ_omit |> 
  ggplot(aes(x = unemployment, y = pv2p, label = year)) + 
  geom_label(aes(family = "Courier")) + 
  geom_smooth(method = "lm", formula = y ~ x, color = "#9999CC") +
  geom_hline(yintercept = 50, lty = 2, color = "#000033") + 
  geom_vline(xintercept = 0.01, lty = 2, color = "#000033") + 
  labs(x = "Second Quarter Unemployment Rate (%)", 
       y = "Popular Vote Share - Incumbent Party (%)", 
       title = "Fig. IV - Q2 Unemployment vs. Vote Share, 2020 Omitted") + 
  my_theme() + 
  theme(plot.title = element_text(size = 18))
```

```{r OLS unemployment}
unemp_with <- lm(pv2p ~ unemployment, 
               data = inc_econ)
unemp_with |> summary()

unemp_without <- lm(pv2p ~ unemployment, 
                         data = inc_econ_omit)
unemp_without |> summary()
```

```{r unemp boxplot}
ggplot(inc_econ, aes(x=unemployment)) +
  geom_boxplot() +
  xlab("Second Quarter Unemployment Rate (%)") +
  geom_label(x=13, y=0.05, label = "1980") +
  ggtitle("Q2 Unemployment Boxplot") +
  my_theme() +
  theme(plot.title = element_text(size = 18))
```

```{r}
gdp_without_res <- resid(gdp_without)
plot(fitted(gdp_without), gdp_without_res)
abline(0,0)
```

# A Tale of Two Incumbents

```{r}

```

# 2024 Prediction

# Sequester 2024 data
```{r sequester 2024}
GDP_new <- fred |> 
  filter(year == 2024 & quarter == 2) |> 
  select(GDP_growth_quarterly)
```

# Predict
```{r predict w/ gdp}
# use q2 gdp for 2024 to predict next pop vote
predict(gdp_without, GDP_new)
```

# Predict uncertainty
```{r predict bounds}
predict(gdp_without, GDP_new, interval = "prediction")
# baseline alpha probably 0.05 (95% confidence)
```